---
title: "in your memory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{in your memory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_in_your_memory.bib
link-citations: true
autosize: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

31 July 2025

## Introduction

Forensic Scene was played live for the first time at Colégio Leão XIII in Itaboraí, RJ, Brazil on 19 August 1994 [@fugazi1994], introduced by Guy with the comment "you're probably going to hate this song but we're going to play it anyway". The song got it's official release on Red Medicine which was released around the 12 May 1995 [@fugazi1995]. It was one of the later songs to debut that would be released on Repeater.

```{r echo=FALSE}
releases_data_input %>%
  filter(release=="red medicine") %>%
  select(song, date) %>%
  rename(debut = date) %>%
  arrange(debut)

```

Fugazi played Forensic Scene at least 201 times with the last recorded version being on the 4 November 2002 in London, England - the final Fugazi show to date. It was the 5th most played song from Red Medicine. In live shows Forensic Scene tended to be used later in the set, it was often followed by Promises (31 times) or by Long Division (21 times).

There are several interviews where Guy Picciotto has commented on the song. In 1995 he only commented in outline:

> I don’t normally discuss what songs are actually specifically about. I feel like that the work that was put into creating the song is the expression and that people who listen to the song, if they’re interested in the song, create their own interpretations. I feel like, if it doesn’t make any sense, then I didn’t succeed when I wrote the song. I’m willing to accept that as a failure, but in general… Sometimes I’ll discuss a song like Cassavetes on In on the Kill Taker. I can say that that’s a song about the filmmaker John Cassavetes, because it seems to be fairly obvious, but Forensic scene… It’s like a less… It’s a more personal song. You can’t say that it’s about a specific person or something like that. It’s more about a situation. And I think it’s up to people to figure it out for themselves. [@mackaye1995]

In a 2025 interview with Brian Gathy for the End on End podcast he went into more detail:

> Brendan actually wrote the verse chords and I wrote the chorus chords... the verse chords seem simple but it's actually a very complicated thing. Brendan writes some very complicated guitar shit. His chordings are so wild and it's one of my favourite things that he ever wrote. The pleasure of learning to play that and then figuring out how to sing to it and then coming up with some big chorus ... it's one of my favourite things that Brendan and I worked on together, and it really came from those initial chordings that Brendan came up with, that are so interesting." [@picciotto2025]

> It was a very sad moment that produced that song, I don't really have anything more specific to say to it than that, ... it was kind of like that forensic analysis post-fact of the horrible thing that had just happened, that's kind of like what it was." [@picciotto2025]

The first time the song was recorded live, in August 1994, it was fully formed with both lyrics and music being more or less identical to the studio version that would be recorded in early 1995. This strongly suggests it had been written, rehearsed and possibly performed live some time earlier. The 1993 Japan / Australia / New Zealand Tour ended in Honolulu, Hawaii on the 14 December 1993 and the next show after that was Brendan's Going Away Party at Pirate House on 3 June 1994 (no recording exists). After that Brendan went to live in Seattle for several months, which means Forensic Scene was probably written sometime between December 1993 and May 1994. What happened in that time period that could be described as a 'horrible thing' and might have prompted the composition of this song? One event in particular comes to mind. To this day I remember where I was when I heard the news, which even on the first telling involved some forensic details - it stuck in my mind that they had to identify him from his fingerprints.

> NIRVANA SINGER FOUND DEAD By Richard Harrington and Richard Lei. April 9, 1994. Kurt Cobain, who led the grunge band Nirvana to superstardom yet reviled his own success, was found shot in the head in his Seattle home yesterday, an apparent suicide. [@harrington1994]

There are other possible interpretations. Several discussions of the song have characterized it as being about the break-up of a relationship [@wright] and [@gathy]. Both of these discussions were prior to Guy Picciotto's most recent remarks, which strongly suggest that it is about something much more unpleasant than a break-up.

Patrick Foster's 1995 review of Red Medicine linked the song to the O.J. Simpson trial but this is impossible because the trial took place in 1995 and the song was already being performed live in August 1994.

> “Forensic Scene” is driven by disgust at the circuslike O.J. Simpson trial, its singsong, nearly Lennonesque vocal fraught with hold-release momentum. [@foster1995]

## Title

The American Academy of Forensic Sciences has this to say about the word 'forensic':

> The word forensic comes from the Latin word forensis: public, to the forum or public discussion; argumentative, rhetorical, belonging to debate or discussion. A relevant, modern definition of forensic is: relating to, used in, or suitable to a court of law. Any science used for the purposes of the law is a forensic science. ... Forensic scientists may be involved anytime an objective, scientific analysis is needed to find the truth and to seek justice in a legal proceeding. [@aafs2022]

'Scene' has several possible meanings [@merriam-webster2025]:

-   one of the subdivisions of a play

-   a stage setting, or a real / imaginary view suggesting a stage setting

-   the place of an occurrence or action

-   an exhibition of anger or indecorous behavior

-   a sphere of activity or situation

## Lyrics

The lyrics from the studio recording will be referred to here [@fugazi1995] - there are some variations in live recordings. The lyric sheet uses all lower case with no line breaks, here I'm going to use lower case and line breaks based on how it is sung.

Notice how many of the lines are couplets and some of them rhyme although the rhymes seem forced at times: gentle / sentimental; without / now; would / wound; lay / anyway; unclean / scene; instant / distance.

Guy's comment that "It’s a more personal song" [@mackaye1995] suggests that the lyrics should be interpreted as being sung from a first-person point of view.

### in your memory not so gentle

"in your memory, not so gentle": something not so gentle in memory of someone.

### perversify slash sentimental

If 'versify' means to turn something into verse, then 'perversify' would mean to make something perverse.

'perversify slash sentimental' suggests being by turns perverse and sentimental about something.

### caught between and caught without

Unexpectedly in a difficult or uncomfortable situation.

Words are omitted leaving the listener to fill in the blanks.

Several sayings come to mind:

-   caught between a rock and a hard place

-   caught without an umbrella

### i'm a failure not your failure now

"I'm a failure, not your failure now" - now it's me, not you anymore.

Taking responsibility for a bad situation or for having done something wrong. Expresses humility and perhaps also regret.

### in and out just like a knife would

Something that is not a knife that penetrates like a knife - maybe a needle.

If it is a needle this could refer to drug use.

### tax the flesh and leave a cheap wound

Unusual usage of tax. The only similar usage I could find was in an academic paper on the etymology of the word 'dismal', which suggests a possible connection to some religious practices such as fasting and flagellation:

> in our sentence, desmar las carns, the verb desmar is used to refer, not to time, but to the flesh... 'to decimate, to diminish'. "now is the time to 'tax' the flesh." The phrase decimare carnem is but a variant of the phrase exuere carnem which, according to Saint Augustine, expresses the duty of the christian who wants to follow Christ. [@spitzer1942]

'Cheap wound' implies that it is something small, not a serious injury, which seems consistent with the idea of a needle being used for injections.

### fucked and fucking love and lay

'fucked and fucking, love and lay' - brings to mind press coverage of celebrities, in particular their misfortunes and their sexual relationships, which may or may not involve love.

### congenitally fractured anyway

Congenital in this context would mean 'existing at or dating from birth', 'an essential characteristic', or 'being such by nature' [@merriam-webster2025a].

A continuation of the previous line, alluding to commentary and speculation about other people's lives and what made them do what they did.

### unclipped unclean

Unclipped nails, dirt or blood under nails could be part of a forensic investigation in order to establish responsibilites.

### this forensic scene's

Forensic seems to be used here in the sense of people trying to figure out what happened, why it happened, and who is to blame.

Most if not all of the meanings of 'scene' seem relevant here. Scene as in a place where something happened, a community of bands and musicians, and an actual or imagined visual perspective.

### all played out

'played out' means exhausted, spent, finished. 'played out' also resonates with one of the meanings of scene - a scene of a play, perhaps alluding to something dramatic about the subject of the song.

### the defense rests

In a trial the defense rests when they have no more evidence to present at that stage. This implies that the process is moving towards some sort of closure.

### and sorry's just a no shit sherlock mouth talk con job

Alludes to a situation that is so serious that an apology is inadequate and perhaps inappropriate

'no shit sherlock', meaning something that is obvious and does not need to be said, ironically references Sherlock Holmes, Arthur Conan Doyle's famous detective character who made use of deductive reasoning to infer the truth behind the observed facts of his cases.

'mouth talk con job' indicates insincerity or deception.

### from your memory to this instant

From thinking about someone in the past to the present moment.

### worked so hard for working distance

This alludes to being too close to something, and perhaps trying to get more distance in order to get a better perspective.

'working distance' in photography refers to the distance between the lens and the object being photographed [@carnathan2014]. The greater the working distance, the easier things are in terms of lighting and composition.

### like a mouth too late to shut

This is a metaphor for having done or said something wrong, implies some degree of regret.

### i'm a failure not your failure now

Outside of the chorus, this is the only line that is repeats, which gives it emphasis.

## Live variations

## Press Cuttings

## Discussion

## Acknowledgements

## References
