---
title: "au clair de la lune"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{au clair de la lune}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_au_clair_de_la_lune.bib
link-citations: true
autosize: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

11 November 2024

## Introduction

Two Beats Off was debuted as an instrumental at the 40 Watt in Athens, Georgia on 3 May 1989. The first rendition with lyrics was at the 9:30 Club in Washington DC on 20 July 1989. Therefore, the initial version of lyrics and vocal part was probably finished in late spring or early summer 1989, most likely in Washington DC after the end of the spring tour. The 9:30 Club show seems to be the source of the photo by Jim Saah that featured on the cover of Repeater, the album on which Two Beats Off got it's official release on 19 April 1990. Fugazi played Two Beats Off at least 371 times with the last recorded version being on the 6 July 2001 in Victoria, BC, Canada. The opening part of Two Beats Off often featured vocal samples of other songs, for instance:

-   Summertime by George Gerschwin (1934)

-   You've Really Got a Hold On Me by Smokey Robinson and The Miracles (1963)

-   The Place I Love by The Jam (1978)

One rendition in particular caught my attention more than any of the others.

> I'm sure most of you know this is a benefit, and for a very important cause indeed. In Washington DC we have similar problems with youth care, everybody's children think other children are crazy apparently, and they put them away in little psychiatric prisons where they are mistreated, abused, molested... So, anyway, I think it is a rather worthy cause. - Ian MacKaye, 7 November 1990 [@fugazi1990]

On 7 November 1990 Fugazi and The Ex played a benefit for Steunpunt Zetten (support point Zetten) at Paradiso in Amsterdam, Netherlands. Steunpunt Zetten was a support group and activist group working on behalf of young people who complained that they suffered abuse of various kinds from a psychiatrist in charge of the residential youth care institution (Heldringstichting) in Zetten. Complaints were originally published in 1974 but it took until 1990 for the psychiatrist Theo Finkensieper to be investigated, charged and sentenced to 6 years in prison. He was sentenced on 30 May 1990. Finkensieper's appeal was heard in October 1990 and rejected. 'When Steunpunt Zetten sought publicity from 24 April 1990, other reports of sexual abuse within youth care institutions came in almost immediately, such as in the youth care institution De Dreef in Wapenveld, where there was sexual abuse of boys. Ultimately, seven employees were also convicted here in 1993' [@vanderlans2012].

Guy Picciotto sang something in French as part of an extended introduction to Two Beats Off [@fugazi1990].

> Au clair de la lune
>
> Je suis tombé dans l'eau
>
> Ça se fait des bulles
>
> C'est très rigolo
>
> Voilà mon amour
>
> Avec des ciseaux
>
> Elle va couper mon coeur
>
> En trois mille morceaux

This is a modified version of a rude nursery rhyme [@bouteloup2003] that is sang to the tune of an older, traditional 18th century folk song. The folk song has many innuendos or 'sexual undertones' and is the subject of the oldest known audio recording (1860) [@wikipediacontributors2024]. Compared to the rude nursery rhyme, Guy mainly changed the second verse, what he sings is "Here is my love / with scissors / she is going to cut my heart / into three thousand pieces", whereas in the nursery rhyme it translates roughly as "My grandmother is coming / with scissors / she cuts my buttocks / in three thousand pieces" [@bouteloup2003]. Guy Picciotto wasn't the first punk rocker to have made use of "au clare de la lune" - Siouxsie Sioux sang the first verse or two (in French) in 'The Lord's Prayer' (1979) [@siouxsieandthebanshees1979].

The extended introduction that was added to Two Beats Off made me wonder why this was done, and in particular it got me thinking that maybe there was a connection between Two Beats Off and the nature of this particular benefit concert. It wouldn't be the first time that something special had been performed by Fugazi at a benefit concert - Guy commented in an interview [@picciotto2021] that 'Give Me The Cure' was written for and performed for the first time at a benefit for AIDS research at Columbia University in New York [@fugazi1988].

## Lyrics

Here the lyrics of Two Beats Off will be used as headings and commented in detail. The lyrics from the studio recording will be referred to here [@fugazi1990a] - there are some variations in live recordings.

### I cut my nails to the quick

Dirt. Fear of getting caught. Evasion. Pain.

### But still I was caught with my hands in the till

Stealing (or some other crime) in the workplace.

### Red-handed

Guilty of crime.

### Give me something, give me anything

Begging, suggests the subject is destitute or chronically deprived of something.

### The threat of everything is when it becomes nothing at all

Fear of losing everything.

### My fingers reaching, the trophy swelling

Trophies are normally won or captured through sport, hunting or war [@cambridgeuniversitypress2024]. All of these involve taking through force or some other form of exertion something which is not freely given. The way the word trophy is used here as part of a sexual innuendo hints at abuse.

### That's when desire trips me up...

Someone who tries to be good but gets led astray by their desire. Similar language is used in the Bible.

> All of us get tripped up in many ways. Suppose someone is never wrong in what they say. Then they are perfect. They are able to keep their whole body under control. [@james3:1996]

### Got a new technique, money lets the pieces fit where they fall

'Got a new technique' speaks to recent developments, perhaps newsworthy events at the time the song was written. Scandals and how people or organisations with money and power adapted to manage them.

Something broken but not cleaned up or fixed, just normalised and made to look like there is nothing wrong. "Move along, there is nothing to see here".

In particular, this bit makes me think of pedophile priests and how they were dealt with by the church. In 1990 there was a TV documentary drama film called 'Judgement' about the case of a pedophile priest in Louisiana that first made headlines in 1985 [@chicagotribune1990].

### Privilege - it sanctions everything

Sanction here is used in the sense of 'to formally give permission for something' [@cambridgeuniversitypress2024a].

In many cases church authorities tried to deal with cases of abuse as an internal matter, not reporting the alleged crimes to the police and moving the offenders to other locations where they could continue abusing.

> The release of about 6,000 pages of documents provided a grim backstage look at the scandal, graphically detailing the patterns of serial abuse by dozens of priests who were systematically rotated to new assignments as church officials kept criminal behavior secret from civil authority. [@editorialboard2013]

### Security - a net under it all

The image of a net seems significant, it brings to mind a tightrope walker, trapeze artist or someone else doing something risky, who risks falling from a height - possibly falling from grace.

### Once there was a body to see, now just the biggest mouth to feed of them all

The subject was young and perhaps attractive once, but not anymore - now they still have insatiable desire but are undesirable.

## Discussion

I had literally heard this song hundreds of times and it never made much sense to me until the "au clair de la lune" introduction grabbed my attention and made me wonder why this had been added as an introduction. The benefit for victims of sexual abuse by a psychiatrist at a children's home made me wonder if Two Beats Off might be about similar cases of abuse and seen from this point of view the meaning of the lyrics seemed to become much clearer.

It is notable how Two Beats Off is sung in the first person as if the singer is playing a role in a film or a work of theatre. It isn't the only Fugazi song that uses this device of playing a role, the other one that immediately comes to mind is Suggestion, where the use of a female point of view alerts the singer is representing someone else. Another one is perhaps Give Me The Cure. In the case of Two Beats Off the effect is more subtle and it took me a long time to notice it.

I strongly suspect this song was written mainly about sexual abuse by priests and how these crimes were covered up by the church. This had already become a scandal in the US at the time the song was written, although the scandal grew much bigger over the following years and broadened in scope to other countries around the world.

However, much of the content can be readily adapted to deal with other types of abuse in the context of institutions, as was done in the case of the 7 November 1990 benefit show at the Paradiso in Amsterdam, when the special extended introduction made Two Beats Off a centrepiece of the show. Two Beats Off was already an established song in the Fugazi repertoire at this point, so while it wasn't written for this show it seems likely that the introduction was added as a way of making it more specific to the occasion.

"Au clair de la lune" from the earliest versions in the 18th century seems to have hidden sexual innuendos but nevertheless it is considered a song for children. The rude nursery rhyme that is sung to the same tune as the original juxtaposes childish activities with violent punishment. Guy's version takes the rude nursery rhyme and makes it more romantic, or at least sexual, but no less violent - here comes my lover, with scissors, to cut my heart in three thousand pieces.

A few renditions started with these lines, which go well with "the biggest mouth to feed of them all".

> salad is the first course
> 
> what should I choose for the main? 
> 
> salad is the first course
> 
> what should I choose for the main? 
> 
> [19900908 Unitas, Wageningen, Netherlands]

Some renditions have lines that sound more explicit: 

> you want it?
>
> you've got to pick it up
>
> then take it
>
> up against the wall
>
> [19900907 Nighttown, Rotterdam, Netherlands]

At the Paradiso show Guy also made significant changes to the lyrics of the song, making it into more of a protest song.

> you want it?
>
> you've got to pick it up
>
> then take it
>
> out into the streets
>
> the fingers reaching
>
> your heart is breaking
>
> my friends
>
> take it to the streets
>
> [@fugazi1990]

Two Beats Off was previously featured on the Alphabetical Fugazi podcast [@wright2022], where the discussion was quite different.

## Conclusion

To me, Two Beats Off seems to be a song about sexual abuse. I reached this conclusion after having listened to the song hundreds of times so it was not too obvious. It took a few significant clues for me to see it from the right point of view but once I did the lyrics seemed to become quite clear. Looking at it now, I wonder what took me so long. This makes me wonder whether there are other Fugazi songs that are like this, like puzzles waiting to be solved.

## References
