---
title: "Au Clair de la Lune"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Au Clair de la Lune}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_au_clair_de_la_lune.bib
link-citations: true
autosize: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

8 December 2024

## Introduction

Two Beats Off was launched as an instrumental at the 40 Watt in Athens, Georgia on 3 May 1989. The first rendition with vocals was at the 9:30 Club in Washington DC on 20 July 1989. The song got it's official release on Repeater which was released around the 1 March 1990 [@fugazi1990e]. Fugazi played Two Beats Off at least 371 times with the last recorded version being on the 6 July 2001 in Victoria, BC, Canada.

## Title

I only have speculation to offer here. My guess is that 'two beats off' was a working title that became permanent. Like many Fugazi songs, 'two beats off' started off as an instrumental, and it seems likely that the title is a reference to something about the music. 'Two beats off' could imply that something about the music falls out of step with the preceding rhythm. If that is right it is probably the five heavy chords that come after 'caught with my hand in the till' and before 'red handed'. You know, the bit that goes DAA DUU DEE DII DUU. It's one of many highlights of the song.

## Lyrics

The lyrics from the studio recording will be referred to here [@fugazi1990a] - there are some variations in live recordings. The lyric sheet uses all caps so capitalization is not clear. Here I'm going with all lower case.

### I cut my nails to the quick

Dirt. Fear of getting caught. Evasion. Pain.

### but still I was caught with my hand in the till

This sounds confessional. Stealing (or some other crime) in the workplace.

### red-handed.

Caught in the act.

### give me something, give me anything

Begging, suggests the subject is destitute or chronically deprived of something.

### the threat of everything is when it becomes nothing at all

Fear of losing everything.

### fingers reaching, trophy swelling

Trophies are normally won or captured through sport, hunting or war [@cambridgeuniversitypress2024]. All of these involve taking through force or some other form of exertion something which is not freely given. The way the word trophy is used here as part of a sexual innuendo hints at abuse.

### that's when desire trips me up....

Someone who tries to be good but gets led astray by their desire. Similar language is used in the Bible.

> All of us get tripped up in many ways. Suppose someone is never wrong in what they say. Then they are perfect. They are able to keep their whole body under control. [@james3:1996]

### got a new technique money lets the pieces fit where they fall.

'Got a new technique' speaks to recent developments, perhaps newsworthy events at the time the song was written. Scandals and how people or organisations with money and power adapted to manage them.

Pieces could be pieces of a jigsaw puzzle, pieces of a game such as chess, or pieces of something broken.

-   'pieces fit' brings to mind a jigsaw puzzle where pieces should only fit in one location and nowhere else

-   chess pieces represent a strict hierarchy that alludes to a traditional social hierarchy with the king being the most important piece, then the queen, the rooks, the bishops, the knights, and the pawns. The reference to money here implies power and hierarchy.

<!-- -->

-   something broken but not cleaned up or fixed, just normalised and made to look like there is nothing wrong.

In particular, this bit makes me think of how cases of sexual abuse were dealt with by church authorities.

> Loyola University of Chicago psychologist Eugene Kennedy, a former priest and writer on church affairs, faults bishops for taking "advice of lawyers on issues that cannot be resolved merely by making the church legally defensible.
>
> **It's the sweating surface of a culture that is corrupting. The church has failed to examine conflicts about human sexuality which throb within it.**
>
> Cover-ups are impossible; where there is darkness we need light."
>
> ...overall, a cover-up pattern persists. Too often, bishops have simply recycled accused priests to a new parish, while the afflicted families were held at arm's length. Because judges have sealed the records of many civil settlements, there is no way to know exactly how much has been paid. But there are clear signs of heavy losses. In a Minneapolis case, the plaintiff's sister testified that he had rejected a \$1.5-million offer. The plaintiff's attorney, Jeffrey Anderson of St. Paul, says he has 30 suits pending for other clients, involving 16 priests in seven states. One Louisiana diocese and its insurers have paid \$15 million to victims of a parish priest. In all at least \$30 million is known to have been paid nationwide.
>
> \- The Washington Post, 1989-09-16. [@berry1989]

In 1990 there was a TV documentary drama film called 'Judgement' about the case of a pedophile priest in Louisiana that first made headlines in 1985 [@chicagotribune1990].

> The recent sexual abuse scandal in the Catholic Church has put the enforceability of confidentiality agreements in settlement under the microscope. For years, the Church used confidential settlements to silence abuse victims. Although these agreements protected the identity of the victim, they also concealed the identities of the priests who often continued to serve at their parishes or other ministries. For example, in 1997, the Roman Catholic Diocese of Albany paid a confidential settlement of just under one million dollars to a man who alleged that "he had been sexually abused for six years" by a priest "who regularly plied him with drugs and alcohol." The settlement fell just below the diocese's one million dollar ceiling; above that amount it must seek the consent of its finance counsel, an eight-member board composed of the diocesan bishop and seven lay people, designed to provide accountability for the diocese's decisions. Thus, the settlement was entirely insulated from any public scrutiny, allowing the abusive priest to remain anonymous and perhaps continue in his position. [@philp2002]

### privilege - it sanctions everything.

Sanction here is used in the sense of 'to formally give permission for something' [@cambridgeuniversitypress2024a].

In many cases church authorities tried to deal with cases of abuse as an internal matter, not reporting the alleged crimes to the police and moving the offenders to other locations where they could continue abusing.

> The release of about 6,000 pages of documents provided a grim backstage look at the scandal, graphically detailing the patterns of serial abuse by dozens of priests who were systematically rotated to new assignments as church officials kept criminal behavior secret from civil authority. [@editorialboard2013]

### security - a net under it all.

The image of a net seems significant, it brings to mind a tightrope walker, trapeze artist or someone else doing something risky, who risks falling from a height - possibly falling from grace.

### once there was a body to see, now just the biggest mouth to feed of them all

The subject was young and perhaps attractive once, but not anymore - now they still have desire but are undesirable. This line is not on the lyrics sheet but was confirmed by Guy Picciotto in correspondence that was reported in the Alphabetical Fugazi podcast [@wright2022].

## Live variations

Two Beats Off probably has more live variations than any other Fugazi song, at least in terms of the use of samples of other songs and variations to the main lyrics of the song. Some of this will be due to practical considerations:

> you'd be looking around the stage going like 'oh no, Ian's broken a string'
>
> so is there a song like Two Beats Off where I can play the opening chord for like 5 minutes
>
> and Brendan and I, we can vamp while Ian gets his guitar sorted out
>
> you know, a lot of it was based on practical considerations and how songs began
>
> \- Guy Picciotto [@mackaye2022]

On several occasions the intro to Two Beats Off contained parts of several other songs threaded together. To reduce repetition of the same lines over and over again, the summary of each rendition will focus on the parts that are different, omitting lines that are the same as elsewhere.

### FLS0105 Athens, GA USA 1989-05-03

The music of Two Beats Off was performed as an instrumental, and it seems that at this time the song did not have a title, it was introduced by Ian MacKaye as "and now, part two of this great instrumental" (part one was Joe #1) [@fugazi1989]. This rendition is an important starting point for a detailed look at the song because the absence of vocal part and lyrics gives clues about when these might have been written.

### FLS0133 Washington, DC USA 1989-07-20

By the time of this show at the 9:30 Club, Two Beats Off had been named and had vocals with an initial set of lyrics. The sound-quality is rated 'poor' and there are parts where something is being sung that I couldn't make out, but from what I could hear, this initial version had only a limited set of lyrics and used a lot of repetition.

> I cut my nails to the quick
>
> but still I was caught with my hand in the till
>
> red-handed x 2
>
> you want it? you gotta pick it up x 3
>
> you gotta pick it up
>
> (inaudible)
>
> I want it. I gotta pick it up x 2
>
> I want it
>
> (inaudible)
>
> up against it all
>
> (inaudible)
>
> I gotta pick it up
>
> I want it
>
> up against it all [@fugazi1989a]

### FLS0148 Hoboken, NJ USA 1989-10-08

If the information on Wikipedia is correct, recording for Repeater finished in September 1989 [@wikipediacontributors2024b]. If that is right the lyrics used on the album version had already been written and sung but were not used in their entirety on this occasion. This rendition at Maxwell's featured lyrics close to the album version, but not the same. In particular, the lines about something, anything, everything and security / privilege were omitted. 'Body sweating' appears as an alternative to 'trophy swelling'.

> fingers reaching
>
> body sweating [@fugazi1989b]

### FLS0151 Amsterdam, Netherlands 1989-11-03

This rendition features lyrics that are very close to the album version, but still with a few variations, here for instance there is taking instead of 'give me':

> taking
>
> taking everything [@fugazi1989c]

This performance also features love instead of money:

> I've got a new technique
>
> love will make the pieces fit where they fall [@fugazi1989c]

What might be a new technique is to have lots of interchangeable parts and vary them with every rendition - it is hard to find any two live renditions of this song that are the same, and it is also a challenge to find a live version with exactly the same lyrics as the album version.

### FLS0224 Denver, CO, USA, 1990-06-08

One of the first renditions featuring a sample of 'Summertime' by Gershwin [@gershwin1935] as part of the introduction.

> summertime and the living is easy
>
> summertime and the living is good
>
> your daddy's rich and your ma' is good looking [@fugazi1990f]

Early June might have felt like summer already in Denver although it would have been a few weeks ahead of the summer solstice. 'Summertime' would appear many more times as part of the intro to Two Beats Off, with variations to the wording.

### FLS0251 Wageningen, Netherlands 1990-09-08

This performance at Unitas features the following introduction:

> salad is the first course
>
> what shall I choose for the main?
>
> salad is the first course
>
> what shall I choose for the main? [@fugazi1990c]

It also features some variations to other lines:

> I want a body to be seen
>
> not just the biggest mouth to feed of them all [@fugazi1990c]

### FLS0272 Linkoping, Sweden 1990-10-02

> Swing low, sweet chariot, coming for to carry me home [@fugazi1990b].

Unusually, this intro to Two Beats Off is sung by Ian MacKaye. This is from an old gospel song that has been performed by many artists including Louis Armstrong, Screamin' Jay Hawkins, Sam Cooke, B. B. King, and Joan Baez, who sang it at Woodstock in 1969 [@wikipediacontributors2024a].

Following on from this, Guy Picciotto takes up the lead vocals with a few lines that are clearly based on 'The Place I Love' by The Jam [@thejam1978], with a few changes to wording and sequence.

> The place I love is nowhere near here
>
> Not within a mile of those trendy do's
>
> Beautiful moss and colourful flowers
>
> Let's take a stand against the world [@fugazi1990b]

The main part of the song features some lines that did not make it into the album version but feature in many live renditions.

> you want it?
>
> you gotta pick it up
>
> and take it
>
> up against the wall [@fugazi1990b]

The end of the song also features Ian MacKaye, who sings something that sounds like this:

> he got in
>
> and disappeared
>
> he got here
>
> and disappeared [@fugazi1990b]

### FLS0304 Amsterdam, Netherlands 1990-11-07

On 7 November 1990 Fugazi and The Ex played a benefit for Steunpunt Zetten (support point Zetten) at Paradiso in Amsterdam, Netherlands.

> I'm sure most of you know this is a benefit, and for a very important cause indeed. In Washington DC we have similar problems with youth care, everybody's children think other children are crazy apparently, and they put them away in little psychiatric prisons where they are mistreated, abused, molested... So, anyway, I think it is a rather worthy cause. - Ian MacKaye, 7 November 1990 [@fugazi1990]

Steunpunt Zetten was a support group and activist group working on behalf of young people who complained that they suffered abuse of various kinds from a psychiatrist in charge of the residential youth care institution (Heldringstichting) in Zetten. Complaints were originally published in 1974 but it took until 1990 for the psychiatrist Theo Finkensieper to be investigated, charged and sentenced to 6 years in prison. He was sentenced on 30 May 1990. Finkensieper's appeal was heard in October 1990 and rejected. 'When Steunpunt Zetten sought publicity from 24 April 1990, other reports of sexual abuse within youth care institutions came in almost immediately, such as in the youth care institution De Dreef in Wapenveld, where there was sexual abuse of boys. Ultimately, seven employees were also convicted here in 1993' [@vanderlans2012].

Guy Picciotto sang something in French as part of an extended introduction to Two Beats Off.

> Au clair de la lune
>
> Je suis tombé dans l'eau
>
> Ça se fait des bulles
>
> C'est très rigolo
>
> Voilà mon amour
>
> Avec des ciseaux
>
> Elle va couper mon coeur
>
> En trois mille morceaux[@fugazi1990]

This is a modified version of a rude nursery rhyme [@bouteloup2003] that is sang to the tune of an older, traditional 18th century folk song. The folk song has many innuendos or 'sexual undertones' and is the subject of the oldest known audio recording (1860) [@wikipediacontributors2024]. Compared to the rude nursery rhyme, Guy mainly changed the second verse, what he sings is "Here is my love / with scissors / she is going to cut my heart / into three thousand pieces", whereas in the nursery rhyme it translates roughly as "My grandmother is coming / with scissors / she cuts my buttocks / in three thousand pieces" [@bouteloup2003]. Guy Picciotto wasn't the first punk rocker to have made use of "au clare de la lune" - Siouxsie Sioux sang the first verse or two (in French) in 'The Lord's Prayer' (1979) [@siouxsieandthebanshees1979].

On this occasion Guy also made significant changes to the lyrics of the song, making it into more of a protest song.

> you want it?
>
> you've got to pick it up
>
> then take it
>
> out into the streets
>
> the fingers reaching
>
> your heart is breaking
>
> my friends
>
> take it to the streets [@fugazi1990]

The extended introduction that was added to Two Beats Off made me wonder why this was done, and in particular it got me thinking that maybe there was a connection between Two Beats Off and the nature of this particular benefit concert. It wouldn't be the first time that something special had been performed by Fugazi at a benefit concert - Guy commented in an interview [@picciotto2021] that 'Give Me The Cure' was written for and performed for the first time at a benefit for AIDS research at Columbia University in New York [@fugazi1988].

"Au clair de la lune" from the earliest versions in the 18th century seems to have hidden sexual innuendos but nevertheless it is considered a song for children. The rude nursery rhyme that is sung to the same tune as the original juxtaposes childish activities with violent punishment. Guy's version takes the rude nursery rhyme and makes it more romantic, or at least sexual, but no less violent - here comes my lover, with scissors, to cut my heart in three thousand pieces.

### FLS0437 Nottingham, England 1992-05-08

This one features what might have been a completely improvised introduction.

> I'm in love
>
> with the old world
>
> I'm in love
>
> young and handsome too
>
> I'm in love, in love
>
> with the old world
>
> (and hand-) and handsome too [@fugazi1992]

### Unidentified Introductions

There are several other shows with extended introductions to Two Beats Off, I haven't transcribed them and I'm not sure whether they are original compositions, improvisations, samples of other songs, or some combination of these. There may be others, but these are the ones I have on my list.

-   FLS0491 Virginia Beach, VA USA 2/4/1993 [@fugazi1993]

-   FLS0765 Washington, DC USA 1/31/1996 [@fugazi1996]

-   FLS0772 Athens, GA USA 3/26/1996 [@fugazi1996a]

## Press Cuttings

As seen in the summaries of live performances, this song went on evolving after Repeater had been released and almost every performance was different in some way from previous versions.

A series of potentially relevant news stories was published at around the time Two Beats Off was in the process of being written, and the story of one particular person, the DC area priest George Augustus Stallings Jr., seems to incorporate many of the themes of the song.

> Stallings's dealings with young men in his parish were the subject of scrutiny in the late 1980s, when a series of front-page articles in The Washington Post quoted two unnamed former altar boys as saying that he had abused them. Stallings's former pastoral assistant, who was 22 at the time, spoke publicly about having a two-year sexual relationship with him.
>
> Stallings denied all of the allegations and had left the Catholic Church by the time the articles were published in 1989. He was excommunicated in 1990.
>
> \- The Washington Post, 2009-10-14 [@wan2009]

It is not possible to say for sure that this person was a subject of the song but there are many aspects of his story that seem to coincide with parts of the song, including some of the variations that were performed live.

### swing low, sweet chariot

> Stallings, 41, a convert to Catholicism, attracted national attention as pastor of St. Teresa of Avila parish in Southeast Washington by melding the gospel music and preaching and worship styles familiar to blacks with the traditional mass. According to the priest, in his 12 years at St. Teresa's, membership increased from 200 to 2,000.
>
> \- The Washington Post, 1989-06-20. [@hyer1989]

### fingers reaching, trophy swelling that's when desire trips me up....

> The man, whom The Washington Post agreed not to identify, said the first incident occurred in the rectory on a summer afternoon when Stallings, whom he respected and considered "a great role model," asked the 16 year-old to kneel next to him on the carpet and share a loaf of bread. "He started touching my ear. And one thing led to another and we ended up upstairs in his room," the man, now 28, said in an interview Friday. "And, you know, that was the start of the relationship right there, at least the sexual relationship."
>
> "He talked about how the breaking of bread between two people is a very, how do you say it, a very special thing, a very caring thing," the man said. He said he and Stallings had oral sex that day and two or three times a week over the next three or four months. "He was saying things along the line that this was okay and that this was fine, this is something very special between two friends, and that it was nothing to be ashamed of or afraid of . . . . I thought it was wrong because this man was a priest and he was supposed to have this vow of chastity." They had sexual relations only in the church rectory, he said, although sometimes, "after mass, he would pull me aside, like into the side room there. He would kiss me, which I thought was awfully bold."
>
> \- The Washington Post, 1989-09-04 [@sessionsstepp1989]

### take a stand against the world

> The Rev. George A. Stallings Jr., the dissident Catholic priest, went on national television yesterday to say that he is innocent of the allegation that he had sex with a 16-year-old altar boy. "Read my lips: I am innocent," Stallings said on CNN's "Larry King Live." Stallings went on to say, "What else do you want to know?" but declined to answer questions. He would not say if he has ever had sex with a minor or any man. He also would not state his views on pedophilia or the vow of celibacy. The Washington Post reported Monday that the former altar boy, now 28, swore in a signed statement that he repeatedly had sexual relations with the priest in the rectory of St. Teresa of Avila Catholic Church in Southeast Washington in a three- to four-month period in 1977. The man, whom The Post agreed not to identify, said he broke off the relationship and never reported it to authorities. The Post also reported that the Archdiocese of Washington had received a separate complaint three years ago but had not been able to substantiate it. This summer, The Post reported, Cardinal James A. Hickey confronted Stallings about his lifestyle and asked him to seek treatment at a church-owned facility specializing in treating pedophilia. Soon after the meeting with Hickey, Stallings defied church authorities by founding the Imani Temple, an independent black Catholic congregation. He said he was prompted by racism in the church.
>
> \- The Washington Post, 1989-09-06. [@dedman1989]

> Although Hickey couldn't establish whether Stallings was breaking his vow of celibacy, he could see that Stallings was breaking another vow - of obedience. In 1987, for instance, Stallings hired without permission a priest who had been suspended from his order in another city because of allegations that he made sexual advances to teenage boys. It is unclear what disturbed Hickey most about Stallings: the questions about his homosexual activity or his repeated challenges in many areas to the bishop's authority. After years of such challenges, Hickey last summer gave Stallings an ultimatum. In a dramatic confrontation, according to Stallings, Hickey ordered him to seek psychiatric treatment for an "excessive ego" at a church hospital known for treating priests with sexual disorders.
>
> \- The Washington Post, 1990-04-30. [@sessionsstepp1990]

### the place I love

> In the summer of 1985, a Roman Catholic priest named George Augustus Stallings Jr. bought his independence for \$60,000 and named it Augustus Manor. The 80-year-old house, just up the hill from his St. Teresa of Avila Church, didn't look like much. But in the next two years Stallings restored it and adorned it with European and Oriental antiques, a Japanese garden and an Italian marble bathroom. The "jewel of Anacostia," as Washingtonian magazine called Augustus Manor, became a sign of Stallings's success as a parish priest and a preacher of national renown. His congregation was on its way to a tenfold increase in membership, a major renovation of its historic sanctuary and a leading role in the adaptation of Roman Catholicism for black believers. But Augustus Manor was more than a symbol. It was an act of defiance: Cardinal James A. Hickey, the archbishop of Washington, has a firm rule that parish priests live in rectories.
>
> \- The Washington Post, 1990-04-29. [@dedman1990]

### I cut my nails to the quick

> Brown, now 25, said he and Stallings, 42, were lovers for the next two years, during which time he lived in Augustus Manor, the priest's house in Anacostia. For nine months of that time, Stallings employed Brown as a pastoral assistant at St. Teresa of Avila parish. Brown and other employees said that much of the time he was out of the parish on trips with Stallings. Brown and Stallings were consenting adults, but Stallings was sworn to celibacy in a church that considers homosexual activity a sin. Brown said Stallings knew parishioners were whispering about their relationship. "It bothered him, but all people could do was insinuate," Brown said. "He didn't think it would catch up with him."
>
> \- The Washington Post, 1990-04-30. [@sessionsstepp1990]

### but still I got caught with my hand in the till

> Church officials said they were aware, for example, that since 1979 there had been scattered complaints from some parishioners at St. Teresa who questioned Stallings's stewardship of church funds. When Stallings bought Augustus Manor and began renovating it, some parishioners raised questions as to how he could afford it on his \$10,000 salary, according to a church official. Stallings has said he paid for the renovation with his earnings from preaching revivals around the country. Documents reviewed by The Washington Post show that Stallings wrote a \$4,478 check on a St. Teresa account for a cedar fence that surrounds Augustus Manor. Parish lay leaders and archdiocesan officials said recently they did not know of any expenditure of parish funds for the priest's home. A spokesman for Stallings declined to say why Stallings used the parish account or whether he reimbursed St. Teresa.
>
> \- The Washington Post, 1990-04-29. [@dedman1990]

### others

The recorded version of 'Two Beats Off' was written about three years before Sinead O'Connor protested about sexual abuse in the Catholic church by ripping up a picture of the pope on Saturday Night Live.

> The song, a chantlike a cappella protest of racism and other forms of injustice, says in part: "Until the philosophy which holds one race superior and another inferior is finally and permanently discredited and abandoned, everywhere is war. \... Until the color of a man's skin is of no more significance than the color of his eyes, I've got to say war!" It concludes with the line "We have confidence in the victory of good over evil."
> After she intoned those words Saturday night, O'Connor, 26, wearing a long white lace dress, very close-cropped hair, a nose stud and a silver necklace with a Star of David, held up a color photo of the pope, began tearing it and said, "Fight the real enemy." [@carmody1992]

Sinead O'Connor suffered a significant backlash following this protest, being banned, mocked and booed on various occasions.

> Speaking with the New York *Times* in 2021, O'Connor said she had no regrets, though the backlash was overwhelming. "I'm not sorry I did it. It was brilliant. But it was very traumatizing. It was open season on treating me like a crazy bitch." [@gajanan2023]

So, it is possible to protest or speak out about uncomfortable issues and powerful individuals openly but doing so can be costly and risky. Other cases which might have deterred a more open approach include the backlash experienced by The Beatles in the US to John Lennon's "more popular than Jesus" remark, which led to several incidents and the band deciding to not tour any more [@upi1966].

## Discussion

I had literally heard this song hundreds of times and it never made much sense to me until the "au clair de la lune" introduction grabbed my attention and made me wonder why this had been added as an introduction. The benefit for victims of sexual abuse by a psychiatrist at a children's home made me wonder if Two Beats Off might be about similar cases of abuse and seen from this point of view the meaning of the lyrics seemed to become much clearer.

It seems likely that one of the subjects of this song was sexual abuse and how it often gets covered up. Sexual abuse in the Roman Catholic Church had already become a scandal in the US at the time the song was written, although much more evidence emerged over the following years. Unfortunately other institutions that work with children also have cases of sexual abuse, including protestant churches [@denney2023], Boy Scouts of America [@cubellis2015], and others. Therefore, much of the content of this song could be adapted to deal with abuse in other institutions, as seems to have been done in the case of the 7 November 1990 benefit show at the Paradiso in Amsterdam, when the special extended introduction made Two Beats Off a highlight of the show. Two Beats Off was already an established song in the Fugazi repertoire at this point, so while it wasn't written for this show it seems likely that the introduction was added as a way of making it more specific to the occasion.

It is notable how Two Beats Off is sung in the first person as if the singer is playing a role in a film or a work of theater. It isn't the only Fugazi song that uses this device of playing a role, the other ones that immediately comes to mind are Suggestion, which is sung from a female point of view, and Give Me The Cure, which seems to be written from the point of view of someone with HIV/AIDS. In the case of Two Beats Off the effect is more subtle and it took me a long time to notice it.

It's very hard to reach definite conclusions about this song because it doesn't give us names, or dates, or locations, there is nothing that can be used to pin it down to some definite meaning or subject, and that was probably intentional on behalf of the songwriter. This doesn't seem to be one of Guy's people songs like Cassavettes or Dear Justice Letter that openly references a specific person - these tended to be about people that Guy admired or valued. This song might be about a specific person but that person is not identifiable - this is not a tribute song.

There are a lot of hints and clues, both in the lyrics themselves and considering the context of the time when the song was written, some of which I have explored in detail. The news stories in the Press Cuttings section may or may not have been used as source material. The number of coincidences with parts of the song make it seem likely that there is a connection but it is impossible to be sure.

The challenge of reaching definite conclusions about the lyrics reminds me of the saying 'there's no smoke without fire' that is used in cases where there is no hard evidence but such a quantity of rumors, speculatation and anecdotal evidence that it seems unlikely that these are all baseless. This seems fitting if the song is about what I think it is about.

## Acknowledgements

Thanks to everyone from 'This is Not a Fugazi Appreciation Group' [@variousauthors2024] who helped transcribe parts of songs that I could not understand by myself, in particular Laure Bertrand, Gunter Habets and Mat Sharp for help with the French.

## References
