---
title: "F/G"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{F/G}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_F_G.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

8 September 2024

## Introduction

Between the 15 of November 2021 and the 4 of February 2024 I listened to all the shows that were available on the Fugazi Live Series site and whatever additional shows I could find - that was 915 Fugazi shows in 812 days [@mitrani2024b]. When I finished the Fugazi Live Series I soon started listening to the only other band I knew of that had a similarly large archive of live recordings - the Grateful Dead.

The Grateful Dead preceded Fugazi by about 20 years, they were young folk musicians when the Beatles changed everything. They started playing electric instruments and their early experiences as house band for the acid tests in San Francisco informed their musical style which made extensive use of improvisation.

Fugazi are considered 'post-hardcore' in that their background was in hardcore punk but they deliberately made a different type of music that was slower and combined influences from several musical styles outside of punk. Fugazi were already a successful band on the independent record label Dischord Records when in 1991 Nirvana released Nevermind and a lot of things changed.

## Opposites

In some ways Fugazi and the Grateful Dead seem like polar opposites:

-   Ian MacKaye wrote a series of songs for his previous bands that were critical of drug users: Deadhead [@theteenidles1980], Straight Edge [@minorthreat1981] and Out of Step [@minorthreat1983]. Fugazi generally refrained from using alcohol, tobacco and drugs, while members of the Grateful Dead used alcohol, tobacco, cannabis, LSD, cocaine and heroin.

-   Grateful Dead songs often feature long improvised guitar solos whereas Fugazi songs don't have guitar solos at all.

-   Fugazi refused to play songs from their previous bands and they did not play cover versions (with a one-off exception when they covered a Bad Brains song), whereas Grateful Dead shows over the years included about 190 cover versions as well as 151 original songs.

-   Fugazi songs were often political while Grateful Dead songs tended to tell stories, they tried to avoid mixing music and politics.

> The Dead's attitude about politics would stay consistent, if unique. They probably did more benefits than any band ever, and frequently for explicitly political groups, but they'd sign nothing. They put their time where their beliefs were, but not their mouths. It was a policy designed to make "Grateful Dead" something separate, above life's daily shuffle. "It's our responsibility to keep ourselves free of those connotations," said Garcia years later. "I want the Grateful Dead experience to be one of those things that doesn't have a hook. We're all very antiauthoritarian. There's nothing that we believe so uniformly and so totally that we could use the Grateful Dead to advertise it." [@mcnally2003]

## Quotes

### Jerry Garcia on the Grateful Dead's Performance at the Monterey Pop Festival in 1967

> [it] "was one of the classic bad scenes for us." Garcia described it this way for a television news interviewer: "We came onstage just after The Who finished smashing their equipment onstage for the first time in America. This is aahhh, the audience is devastated. The Who were beautifully theatrical, there's clouds of smoke and explosions and they're clearing away the debris. And so we came out and played our little act --- ding ching ding ding --- and then Jimi Hendrix comes on after us. We were erased from existence." [@livernois2019]

### Jerry Garcia on Punk

> DJ: "You're not into any disco or New Wave or anything?"
>
> JG: "Of those things, of those choices, the New Wave\... the whole punk rock trip, I like it best. I like it best, I think. I like it because it has some energy, it has some spirit."
>
> JK: "There's the thing of, new music, people are tyring to play [crosstalk] there's thing, in San Francisco, 'cause that's where I was then, when everybody was just learning to play, there was a thing that was missing \--"
>
> JG: "The thing of people tryin' real hard.
>
> JK: "Tryin' to figure out how to make it work."
>
> JG: "Yeah, yeah. It's got the spirit, y'know, that's the thing. The nature of the spirit, if you had to describe the spirit, you could describe the spirit as being a little bit edgy, y'know, maybe sort of a Violent spirit, y'know, sort of scaly and weird, maybe, but it's a spirit, and almost anything else lacks spirit of any sort." [@garcia1978]

> "It's always nice for the kids to have music. Punk music at it's finest was music that the kids could dig, and it was thoroughly obnoxious to adults. It had that great quality of being able to communicate to the kids and give them something to talk about, something to love, and also alienating everybody else. For me, it's the thing of, when you see young musicians, playing in a way, three chord rock 'n' roll, you know there's only one place for them to go, they're either going to get better, and finally learn how to tune their instruments, and do that kind of stuff, and evolve more, or else they're gonna shed more blood. For me, commitment counts, when you're on stage, and whether you're bleeding over every note, whether you're wounding yourself physically on whatever level, that's real." - Jerry Garcia (1987) [@garcia1987]

### Ian MacKaye on the Grateful Dead

> "I thought it was interesting that they played 2,300-2,600 shows, and I was shocked. I thought they would have played around 5,000 shows, or something way beyond what they actually did."
>
> [interviewer] "Did you ever listen to them?"
>
> "No. But I think they\'re interesting structurally, in retrospect, for instance their business and the way they flexed their muscles in a good way. In the late 80s or early 90s, before Jerry died, they did these stadium shows, and they had their own ticketing system. They sold tickets directly, and at some point during this stadium tour, Ticketmaster didn\'t want them selling directly. They told The Dead that wasn\'t happening anymore, so the band told them, 'Okay, you\'re fired'. And then so Ticketmaster tried to backtrack. At least that\'s what I heard, I\'m not sure if that\'s necessarily true. But anyways, they operated well, they weren\'t bullied by the machine. You get that sort of power by saying 'No'.\" - Ian MacKaye (2013) [@mackaye2013]

## Head to Head

+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Topic                       | Fugazi                                | Grateful Dead                                         |
+:============================+:======================================+:======================================================+
| Status                      | On indefinite hiatus                  | Disbanded                                             |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Hometown                    | Washington, DC                        | Palo Alto, CA                                         |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Period of Activity          | 16 years (1987 - 2002)                | 31 years (1965 - 1995)                                |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Original songs played live  | 94 [@mitrani2024]                     | 151 [@amato2015]                                      |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Cover versions played live  | 1                                     | 190 [@amato2016]                                      |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Live performances           | 1048 [@mitrani2024]                   | 2325 [@gratefulseconds2016]                           |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Benefit shows               | 80 [@klirs2018]                       | 99 [@fatemusic2018]                                   |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Studio albums               | 6                                     | 13                                                    |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Live albums                 | 896 [@dischordrecords2024]            | 167 [@guinnessworldrecords2024]                       |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Live recordings             | 915 [@mitrani2023]                    | 2065 [@relisten2024]                                  |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Members (past and present)  | 4                                     | 13                                                    |
|                             |                                       |                                                       |
|                             | [@dischordrecords2024a]               | [@thegratefuldead2023]                                |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Touring musicians           | 1                                     | 2                                                     |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Main instruments            | Vocals, Guitar, Bass, Drums           | Vocals, Guitar, Keyboards, Bass, Drums.               |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Other instruments           | Trumpet, Clarinet, Soundboard effects | Harmonica, Trumpet, Saxophone,                        |
|                             |                                       |                                                       |
|                             |                                       | Clavichord, Synthesizer, Computer                     |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| USA States played           | 50 [@dischordrecords2024]             | 45 [@golsen2022]                                      |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Countries played            | 36 [@mitrani2024a]                    | 12 [@twelvebarsoftwarellc2024]                        |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Venues played               | 733 [@mitrani2024a]                   | 592 [@twelvebarsoftwarellc2024]                       |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Average concert length      | 1 hour 17 minutes [@mitrani2024]      | 3 hours [@teachrock.org2021]                          |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Length of longest concert   | 2 hours and 3 minutes [@mitrani2024]  | 4 hours 41 minutes [@internetarchivecontributors2014] |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Average song length         | 3 minutes 34 seconds [@mitrani2024a]  | 9 minutes [@teachrock.org2021]                        |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Longest live song rendition | 13 minutes 52 seconds [@mitrani2024]  | 46 minutes 59 seconds [@twelvebarsoftwarellc2024]     |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Average attendance          | 938 [@mitrani2024]                    | 9604 [@gratefulseconds2016]                           |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Largest live audience       | 15,000 [@mitrani2024]                 | 600,000 [@teachrock.org2021]                          |
+-----------------------------+---------------------------------------+-------------------------------------------------------+
| Total live audience         | 983,881 [@mitrani2024]                | 22,330,524 [@gratefulseconds2016]                     |
+-----------------------------+---------------------------------------+-------------------------------------------------------+

## Live Performances

Both Fugazi and the Grateful Dead would play different sets each show with no pre-defined setlist except for perhaps agreeing on which songs to start on.

### Fugazi

> "We would get on stage with no setlist and then it was a freefall from there"
>
> "We worked out this system of hand signals... the one thing we knew was that Ian and I would trade vocals... usually it was like, if I started a song on guitar, then they would just have to be listening, if Brendan had to start it, I had to get him the message" 
>
> "The reason why it's a good idea and the reason why I'm shocked other bands don't do it is because it forces a level of attention on you and I kind of tightrope aspect to it that it doesn't let you relax in the set"
>
> \- Guy Picciotto [@picciotto2019]

### Grateful Dead

> 'We don't make up our sets beforehand. We'd rather work off the tops of our heads than off a piece of paper.' - Jerry Garcia [@gans1985]

> 'Jerry Garcia and Bob Weir planned out only two things before each of the thousands of concerts they played together in the Grateful Dead: The song they would begin with and the song they would end with. Everything else was decided in the moment. "While Jerry was singing, I had plenty of time to think about the next tune that I was going to do," says Weir. "And while I was singing, he had plenty of time to think about what tune he was next going to do. We developed a feel for that."' [@doyle2020]

## Lyrics

Both bands would often use lyrics that were open to interpretation and would be reluctant to explain the meaning of their songs.

### Fugazi

> "I really do feel that my songs are not as inscrutable as they seem, but that is one I can't really defend, maybe it's the exception that proves the rule ... I think a lot of it has to do with engagement but also because I wrote the songs ... I always have this feeling that 'isn't that perfectly clear?' ... a song like let's say 'Last Chance for a Slow Dance' I remember reading something where someone was like 'this is just some more word-salad' ... that song is so personal and I felt so clear ... the part that goes like 'Flare fakes a flower / A burnt-out shower'. Someone was like, just thought that was gobbledigook or something, like to me it's the image of a flare being shot into the sky as a signal for help and it makes the form of a flower in the sky and it makes a burned-out shower of sparks falling and no-body can see it, it seems like such an obvious lonely image, and then people are like 'I really don't know what the fuck that means', it feels like a failure on one hand because it didn't work, but on the other hand I get kind of like 'shit, really? come on!' When I listen to a band I really like I enjoy that engagement, why would anyone go up to Bob Dylan and say like you know 'It's all Over Now Baby Blue' what's that song about? What is that song not about?" - Guy Picciotto. [@mackaye2022]

### Grateful Dead

> 'In "Long Strange Trip," there's a sequence in which Bar-Lev, after years of trying to get to Hunter, is granted an audience backstage after a solo gig. Hunter was always elusive and notoriously averse to discussing the meanings or sources of his compositions (though he was insistent about the songwriter's claim to royalties and credit, as the Internet undermined each). In the film, Hunter recites the lyrics to "Dark Star" ("Dark star crashes, pouring its light into ashes / Reason tatters, the forces tear loose from the axis / Searchlight casting for faults in the clouds of delusion / Shall we go, you and I while we can / Through the transitive nightfall of diamonds?") and says, "What's unclear about that?" Then he kicks Bar-Lev out of his dressing room.'
>
> '"Meaning is often a subterfuge to distract the listener's attention from a writer's lack of multiple resources," Hunter writes. Then he presents an exegesis of his 1975 song "Franklin's Tower," which became a great crowd-pleaser, in Garcia's hands. Hunter cites, among some of his references, Ben Franklin, the Constitution, Pete Seeger, the Bible, E. E. cummings, Bonnie Dobson, an Eastertide anthem called "Roll Away the Stone," and the birth of his son.
>
> "Well, now that you know what I meant by it, it's no great shakes, is it?" Hunter concludes. "Mystery gone, the magician's trick told, the gluttony for 'meaning' temporarily satisfied, one can now take issue with my intent and avoid the song itself."' [@paumgarten2019]

## References
