---
title: "F/G"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{F/G}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_F_G.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

1 September 2024

## Introduction

Between the 15 of November 2021 and the 4 of February 2024 I listened to all the shows that were available on the Fugazi Live Series site and whatever additional shows I could find - that was 915 Fugazi shows in 812 days [@mitrani2024b]. I had become accustomed to listening to a new live show every day and when I finished the Fugazi Live Series I almost immediately started listening to the only other band I knew of that had a similarly large archive of live recordings - the Grateful Dead.

## Facts

+----------------------------+---------------------------------------+---------------------------------------------------+
| Topic                      | Fugazi                                | Grateful Dead                                     |
+:===========================+:======================================+:==================================================+
| Status                     | On indefinite hiatus                  | Disbanded                                         |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Hometown                   | Washington, DC                        | San Francisco, CA                                 |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Period of Activity         | 16 years (1987 - 2002)                | 31 years (1965 - 1995)                            |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Original songs played live | 94 [@mitrani2024]                     | 151 [@amato2015]                                  |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Cover versions played live | 1                                     | 190 [@amato2016]                                  |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Live performances          | 1048 [@mitrani2024]                   | 2325 [@gratefulseconds2016]                       |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Benefit shows              | 80 [@klirs2018]                       | 99 [@fatemusic2018]                               |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Studio albums              | 6                                     | 13                                                |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Live albums                | 896 [@dischordrecords2024]            | 167 [@guinnessworldrecords2024]                   |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Live recordings            | 915 [@mitrani2023]                    | 2065 [@relisten2024]                              |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Members (past and present) | 4                                     | 13                                                |
|                            |                                       |                                                   |
|                            | [@dischordrecords2024a]               | [@thegratefuldead2023]                            |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Touring musicians          | 1                                     | 2                                                 |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Main instruments           | Vocals, Guitar, Bass, Drums           | Vocals, Guitar, Bass, Keyboards, Drums.           |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Other instruments          | Trumpet, Clarinet, Soundboard effects | Harmonica, Trumpet, Saxophone,                    |
|                            |                                       |                                                   |
|                            |                                       | Clavichord, Synthesizer, Computer                 |
+----------------------------+---------------------------------------+---------------------------------------------------+
| USA States played          | 50 [@dischordrecords2024]             | 45 [@golsen2022]                                  |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Countries played           | 36 [@mitrani2024a]                    | 12 [@twelvebarsoftwarellc2024]                    |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Venues played              | 733 [@mitrani2024a]                   | 592 [@twelvebarsoftwarellc2024]                   |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Average concert length     | 1 hour 17 minutes [@mitrani2024]      | 3 hours [@teachrock.org2021]                      |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Length of longest concert  | 2 hours and 3 minutes [@mitrani2024]  | 6 hours [@teachrock.org2021]                      |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Average song length        | 3 minutes 34 seconds [@mitrani2024a]  | 9 minutes [@teachrock.org2021]                    |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Longest live song          | 13 minutes 52 seconds [@mitrani2024]  | 46 minutes 59 seconds [@twelvebarsoftwarellc2024] |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Average attendance         | 938 [@mitrani2024]                    | 9604 [@gratefulseconds2016]                       |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Largest live audience      | 15,000 [@mitrani2024]                 | 600,000 [@teachrock.org2021]                      |
+----------------------------+---------------------------------------+---------------------------------------------------+
| Total live audience        | 983,881 [@mitrani2024]                | 22,330,524 [@gratefulseconds2016]                 |
+----------------------------+---------------------------------------+---------------------------------------------------+

## Live Performances

Both Fugazi and the Grateful Dead would play different sets each show with no pre-defined setlist except for perhaps agreeing on which songs to start on.

### Fugazi

> "We would get on stage with no setlist and then it was a freefall from there"
>
> "We worked out this system of hand signals... the one thing we knew was that Ian and I would trade vocals... usually it was like, if I started a song on guitar, then they would just have to be listening, if Brendan had to start it, I had to get him the message" 
>
> "The reason why it's a good idea and the reason why I'm shocked other bands don't do it is because it forces a level of attention on you and I kind of tightrope aspect to it that it doesn't let you relax in the set"
>
> \- Guy Picciotto [@picciotto2019]

### Grateful Dead

> 'We don't make up our sets beforehand. We'd rather work off the tops of our heads than off a piece of paper.' - Jerry Garcia [@gans1985]

> 'Jerry Garcia and Bob Weir planned out only two things before each of the thousands of concerts they played together in the Grateful Dead: The song they would begin with and the song they would end with. Everything else was decided in the moment. "While Jerry was singing, I had plenty of time to think about the next tune that I was going to do," says Weir. "And while I was singing, he had plenty of time to think about what tune he was next going to do. We developed a feel for that."' [@doyle2020]

## Lyrics

Both bands would often use lyrics that were open to interpretation and would be reluctant to explain the meaning of their songs.

### Fugazi

> "I really do feel that my songs are not as inscrutable as they seem, but that is one I can't really defend, maybe it's the exception that proves the rule ... I think a lot of it has to do with engagement but also because I wrote the songs ... I always have this feeling that 'isn't that perfectly clear?' ... a song like let's say 'Last Chance for a Slow Dance' I remember reading something where someone was like 'this is just some more word-salad' ... that song is so personal and I felt so clear ... the part that goes like 'Flare fakes a flower / A burnt-out shower'. Someone was like, just thought that was gobbledigook or something, like to me it's the image of a flare being shot into the sky as a signal for help and it makes the form of a flower in the sky and it makes a burned-out shower of sparks falling and no-body can see it, it seems like such an obvious lonely image, and then people are like 'I really don't know what the fuck that means', it feels like a failure on one hand because it didn't work, but on the other hand I get kind of like 'shit, really? come on!' When I listen to a band I really like I enjoy that engagement, why would anyone go up to Bob Dylan and say like you know 'It's all Over Now Baby Blue' what's that song about? What is that song not about?" - Guy Picciotto. [@mackaye2022]

### Grateful Dead

> 'In "Long Strange Trip," there's a sequence in which Bar-Lev, after years of trying to get to Hunter, is granted an audience backstage after a solo gig. Hunter was always elusive and notoriously averse to discussing the meanings or sources of his compositions (though he was insistent about the songwriter's claim to royalties and credit, as the Internet undermined each). In the film, Hunter recites the lyrics to "Dark Star" ("Dark star crashes, pouring its light into ashes / Reason tatters, the forces tear loose from the axis / Searchlight casting for faults in the clouds of delusion / Shall we go, you and I while we can / Through the transitive nightfall of diamonds?") and says, "What's unclear about that?" Then he kicks Bar-Lev out of his dressing room.'
>
> '"Meaning is often a subterfuge to distract the listener's attention from a writer's lack of multiple resources," Hunter writes. Then he presents an exegesis of his 1975 song "Franklin's Tower," which became a great crowd-pleaser, in Garcia's hands. Hunter cites, among some of his references, Ben Franklin, the Constitution, Pete Seeger, the Bible, E. E. cummings, Bonnie Dobson, an Eastertide anthem called "Roll Away the Stone," and the birth of his son.
>
> "Well, now that you know what I meant by it, it's no great shakes, is it?" Hunter concludes. "Mystery gone, the magician's trick told, the gluttony for 'meaning' temporarily satisfied, one can now take issue with my intent and avoid the song itself."' [@paumgarten2019]

## References
