---
title: "F/G"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{F/G}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_F_G.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

25 August 2024

## Introduction

This article is a collection of information comparing and contrasting Fugazi and the Grateful Dead.

| Topic                      | Fugazi                                                   | Grateful Dead                                                                                                                                                                                                                                                                                                                                                                                                       |
|---------------|-----------------------------|-----------------------------|
| Status                     | On hiatus                                                | Disbanded                                                                                                                                                                                                                                                                                                                                                                                                           |
| Period of Activity         | 16 years (1987 - 2002)                                   | 31 years (1965 - 1995)                                                                                                                                                                                                                                                                                                                                                                                              |
| Live performances          | 1048                                                     | 2318 [@teachrock.org2021]                                                                                                                                                                                                                                                                                                                                                                                           |
| Studio albums              | 6                                                        | 13                                                                                                                                                                                                                                                                                                                                                                                                                  |
| Live albums                | 896 [@dischordrecords2024]                               | 167 [@guinnessworldrecords2024]                                                                                                                                                                                                                                                                                                                                                                                     |
| Live recordings            | 915 [@mitrani2023]                                       | 2317 [@wkbw7newsbuffalo2017]                                                                                                                                                                                                                                                                                                                                                                                        |
| Members (past and present) | 4 (Brendan Canty, Joe Lally, Ian MacKaye, Guy Picciotto) | 13 (John Perry Barlow, Tom Constanten, Jerry Garcia, Donna Jean Godchaux, Keith Godchaux, Mickey Hart, Robert Hunter, Bill Kreutzmann, Phil Lesh, Ron "Pigpen" McKernan, Brent Mydland, Bob Weir, Vince Welnick). [@thegratefuldead2023]                                                                                                                                                                            |
| USA States played          | 50 [@dischordrecords2024]                                | 45 [@golsen2022]                                                                                                                                                                                                                                                                                                                                                                                                    |
| Countries played           | 36 [@mitrani2024a]                                       | 13 [@wolfman920972019]                                                                                                                                                                                                                                                                                                                                                                                              |
| Live performances          |                                                          | 'No two shows were the same, although many were similar. Even on good nights, they might stink it up for a stretch, and on bad ones they could suddenly catch fire---a trapdoor springs open. Then, there were the weird inimitable gigs, the yellow lobsters. Variation was built into the music. They played their parts as if they were inventing them on the spot, and sometimes they were. ' [@paumgarten2012] |
| Length of longest concert  | 2 hours and 3 minutes [@mitrani2024]                     | 6 hours [@teachrock.org2021]                                                                                                                                                                                                                                                                                                                                                                                        |
| Average song length (live) | 3 minutes 34 seconds                                     | 9 minutes [@teachrock.org2021]                                                                                                                                                                                                                                                                                                                                                                                      |
| Total live audience        | 1 million [@mitrani2024]                                 | 25 million [@teachrock.org2021]                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |
|                            |                                                          |                                                                                                                                                                                                                                                                                                                                                                                                                     |

## References
