---
title: "Head to Head"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Head to Head}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_head_to_head.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

20 September 2024

> Ahistorical\
> You think this shit just dropped right out of the sky\
> \
> - Bulldog Front by Fugazi [@fugazi1988]

## Introductions

The Grateful Dead and Fugazi are in some ways opposites but there are also similarities between the two bands. Let's have a close look at how these two idiosyncratic and pioneering American bands compare.

### Grateful Dead

The Grateful Dead preceded Fugazi by about 20 years, they were young musicians with backgrounds in folk and classical music when the Beatles changed everything. They started playing electric instruments and their early experiences as house band for the acid tests in San Francisco informed their musical style which made extensive use of improvisation.

> We're in the transportation business. We move minds.
>
> \- Mickey Hart [@mcnally2003]

Michael Azerrad wrote about a conversation with Brendan Canty of Fugazi in which they discussed who might be considered 'the great American band', considering and dismissing candidates including Creedence Clearwater Revival, the Beach Boys, Hüsker Dü, The Replacements, REM, The Ramones, and Parliament-Funkadelic. Let's skip to the end:

> And then the conversation finally came around to facing up to the truth.  If you're looking for a band that embodied the American spirit of individualism, who were popular from the '60s through the '90s, who nurtured a vast, multi-generational community around itself, a band that appealed to and drew from north and south, east and west, contained elements of American idioms like country, bluegrass, blues and folk, that not only reflected but affected the spirit of its time and its culture, a band that contributed many iconic songs to the rock lexicon, influenced musicians as disparate as Elvis Costello, Dave Matthews and Animal Collective, a band that appealed to beatniks, hippies and, yes, punks, and did it all on their own terms, well, it's hard to beat the Grateful Dead.\
>     So I'm going to come right out and say it: the Grateful Dead were the Great American Band.
>
> \- Michael Azerrad [@azerrad2010]

The Grateful Dead played the Monterey Pop Festival in 1967, in the unenviable position of following The Who and preceding Jimi Hendrix.

> [it] "was one of the classic bad scenes for us." Garcia described it this way for a television news interviewer: "We came onstage just after The Who finished smashing their equipment onstage for the first time in America. This is aahhh, the audience is devastated. The Who were beautifully theatrical, there's clouds of smoke and explosions and they're clearing away the debris. And so we came out and played our little act --- ding ching ding ding --- and then Jimi Hendrix comes on after us. We were erased from existence." [@livernois2019]

Although he probably does not remember seeing The Grateful Dead, it seems likely that a young Ian MacKaye (who would have been about 5 years old) was in the audience - in interviews he recalls seeing both The Who and Jimi Hendrix at Monterey.

> At some point, I thought I was gonna try to learn guitar because I was very much into Jimi Hendrix and loved rock & roll and I wanted to play guitars and I wanted to break guitars. I had seen the Who at Monterey Pop and Hendrix at Monterey Pop, and it just seemed like 'Wow, I wanna do that.'
>
> \- Ian MacKaye [@mackaye2012]

### Fugazi

Fugazi are considered 'post-hardcore' in that their background was in hardcore punk but they deliberately made a different type of music that was slower and combined influences from several musical styles outside of punk. Fugazi were already a successful band on the independent record label Dischord Records when in 1991 Nirvana released Nevermind.

> For me punk was construction, I'm a construction worker. - Ian MacKaye [@mackaye2017]

## Drugs

Perhaps the most glaring difference between the Grateful Dead and Fugazi is their attitudes to recreational drugs, with members of the Dead having used all sorts of drugs and Fugazi coming out of the Washington, DC music scene that gave rise to the term "straight-edge" [@kendall2024].

### Grateful Dead

Members of the Grateful Dead used drugs including alcohol, tobacco, cannabis, LSD, magic mushrooms, nitrous oxide, cocaine and heroin. They did not all use all of these drugs, they were not always using, and some of them managed to overcome some addictions. However, it seems fair to say that the band consumed quite a lot of drugs.

> Interviewer: When it comes to drugs, I think the public perception of the Dead is that they are into pot and psychedelics --- sort of fun, mind-expansion drugs. Yet Brent died of a cocaine and morphine overdose, and you also had a long struggle with heroin. It seems to run counter to the image of the band.\
> JG: Yeah, well, I don't know. I've been round and round with the drug thing. People are always wanting me to take a stand on drugs, and I can't. To me, it's so relativistic, and it's also very personal. A person's relationship to drugs is like their relationship to sex. I mean, who is standing on such high ground that they can say: 'You're cool. You're not.' For me, in my life, all kinds of drugs have been useful to me, and they have also definitely been a hindrance to me. So, as far as I'm concerned, the results are not in. Psychedelics showed me a whole other universe, hundreds and millions of universes. So that was an incredibly positive experience. But on the other hand, I can't take psychedelics and perform as a professional. I might go out onstage and say, 'Hey, fuck this, I want to go chase butterflies!'
>
> Interviewer: Does anyone in the Dead still take psychedelics?\
> JG: Oh, yeah. We all touch on them here and there. Mushrooms, things like that. It's one of those things where every once in a while you want to blow out the pipes. For me, I just like to know they're available, just because I don't think there's anything else in life apart from a near-death experience that shows you how extensive the mind is. And as far as the drugs that are dead enders, like cocaine and heroin and so forth, if you could figure out how to do them without being strung out on them, or without having them completely dominate your personality... I mean, if drugs are making your decisions for you, they're no fucking good. I can say that unequivocally. If you're far enough into whatever your drug of choice is, then you are a slave to the drug, and the drug isn't doing you any good. That's not a good space to be in.
>
> \- Jerry Garcia [@garcia1991]

The stereotype of a deadhead is someone who takes drugs, but not all deadheads are like this. The Wharf Rats is a support group for 'concert-goers who have chosen to live drug and alcohol free' [@wharfrats2021] that has held meetings at concerts since the late 1970s, using yellow balloons to help people find them [@epstein1990]. The Wharf Rats is named after the Grateful Dead song Wharf Rat which is about a down-and-out alcoholic who hopes of one day recovering:

> But I'll get back on my feet again someday\
> The good lord willing, if he says I may\
> I know that the life I'm living's no good\
> I'll get a new start, live the life I should
>
> \- Wharf Rat by the Grateful Dead, words by Robert Hunter, music by Jerry Garcia [@hunter1971]

The Wharf Rats still exist (there is a facebook group with 11,000 members at the time of writing) and there are similar support groups for fans of several other jam bands, such as the Phellowship for Phish fans [@wikipediacontributors2024].

### Fugazi

About 8 years before Fugazi formed, Ian MacKaye co-authored 'Deadhead' with Mark Sullivan when they were in the Slinkees, a song which later appeared on the 'Minor Disturbance' EP by the Teen Idles, released as Dischord 1 [@mackaye1989].

> Riding that train high on cocaine\
> The music is really lousy, the fans are a pain\
> Troubles behind, troubles ahead\
> The only good deadhead is one that's dead
>
> Deadhead, deadhead, take another toke\
> Deadhead, deadhead, you're a lousy joke\
> Friend of the devil, who you trying to kid?\
> Friends of the devil are dead like Sid
>
> \- 'Deadhead' by the Teen Idles [@theteenidles1980a]

The lyrics have a strong anti-drugs message and make fun of two Grateful Dead songs, 'Casey Jones' and 'Friend of the Devil'. Curiously, the song 'Casey Jones' itself had an anti-drugs message as it was about a locomotive driver who crashed his train while high on cocaine. It was based on a real historical incident where the driver was known to be teetotal - Robert Hunter changed the story to make it a cautionary tale about the perils of drug use [@dodd1998].

Ian MacKaye wrote a series of songs for Minor Threat that were opposed to the use of alcohol, tobacco and recreational drugs, in particular 'Straight Edge' [@minorthreat1981] and 'Out of Step' [@minorthreat1983]. He later clarified that these songs were mainly about his point of view and his personal decisions, they were not intended to tell other people what to do or what not to do.

> When I wrote this song, Straight Edge, I was not trying to create a set of rules. I was trying to talk about my own decisions and my own life and the right to live my life however I want to. Cause I'm the one that's in this body. So, it's only about my right to live however I damn well please. People interpreted it as, like, a code of behaviour. If you adhere to that you belong to like this mass of like, some kind of, whatever... Union of young men! For the most part. Ahm, I'm just not interested in that. Because, in the beginning, I was straight in a world full of kids who smoked pot, did drugs or whatever. So I felt very alone and I wanted to talk about, like, my validity as a human being, even if I didn't do what everybody else said. Ten years later I see... Or fifteen years later now... It's not such of a problem now, maybe in the late eighties it was more of a problem... I see kids who are trying to inflict their opinions on other people, and their opinions are... You shouldn't be doing this, you shouldn't be doing this, you shouldn't be doing this... It's sort of the opposite of what I intended.
>
> \- Ian MacKaye [@karlsson1995]

Fugazi do not let themselves be easily labelled or categorized, and definitely not as 'straight edge'. As pointed out in the documentary Instrument, Fugazi are not monks [@cohen1999] but they seem to have been a relatively clean-living rock band with no known problems of addiction to alcohol, tobacco or recreational drugs.

> Interviewer: Ian Mackaye is considered the inventor of straight edge. Do you also live straight edge? Guy Picciotto: I don't recognize "straight edge" as a label or a movement. That was the name of a song by Minor Threat, and the ideas behind it are still relevant and interesting. What people have made of 'straight edge' - I can no longer identify with that. Back then, I wouldn't have thought that people would define themselves as 'straight edge'. For me, the song is also about not putting any labels on yourself, that everyone is responsible for what they want to do and not do. It is about being your own person. In that sense, I definitely feel like I live by it. I'm not interested in identifying with this or that. I think that's what cops do, that's not our thing!
>
> \- Guy Picciotto

Fugazi tried to play all-ages venues as much as possible, which often meant venues that did not sell alcohol, and they also avoided giving interviews to large magazines that included advertisements for alcohol and tobacco products.

> The band was now routinely committing rock & roll heresy by turning down big interviews with major national music magazines like Rolling Stone, Details, and Spin, partly because they carried tobacco and alcohol advertising. [@azerrad2001]

## Music

### Grateful Dead

Grateful Dead songs often feature long improvised guitar solos, particularly when performed live.

Grateful Dead shows over the years included 151 original songs [@amato2015] as well as about 190 cover versions [@amato2016a].

The Grateful Dead had 5 keyboard players who were full members and 2 keyboard players who were touring members. The names of the keyboard players are used by fans to refer to different eras of the band.

> JG: "For me music is a full range of experience. In music there is room for space, there's room for quietness, for sorrow, violence."
>
> Interviewer: I do agree. We still seem to be talking about different things though.\
> JG: "It's not my desire to say there is only this or that. For me it's a full range of experiences, and within that it includes things like boredom. Sometimes boredom is what is happening in life, that's what it's about sometimes. Sometimes the tension between boredom and discovery is like an interesting thing. The idea of noodling around aimlessly for 15 minutes, and we are notorious for that, but then hitting on some rich vein of something that we may never have got to any other way, and that's the reward. I want there to be a complete vertical experience. I want it to be the full range."
>
> \- Jerry Garcia [@garcia1981]

XX

> My favorite band was always and still is the Grateful Dead, and they always had that element to their set and I really liked that. There\'s an element to their sets where it goes off the anchor and just gets into sound more than a groove or a riff.
>
> - Greg Ginn [@ginn2013]

XX

> I always thought Jerry Garcia was a musician whose talent was under mentioned. Having been lucky enough to have seen him perform a few times, I always walked away thinking he was a complete player, a musician\'s musician. He had the whole thing: voice, chops, great looking axe. He had a perfect instinct, amazing skill and something that only the truly great guitarists ever achieve -- Jerry Garcia had tone. Damn, what a sound he had. There is nothing like it anywhere. That for me, was the best part of his immense talent; the warm, intimate glow that surrounded the notes he played and the way they hung in the air. Sometimes you would almost forget to breathe, you were listening so intently.
>
> \- Henry Rollins [@rollins2012]

### Fugazi

Fugazi songs don't feature guitar solos.

Fugazi never had a keyboard player, although the piano was Ian Mackaye's first instrument and some Fugazi songs were initially composed on the piano.

Fugazi refused to play songs from their previous bands and they did not play cover versions (with a one-off exception when they covered a Bad Brains song).

> If you look at my work, by and large, I take a position or approach on something and I stick with that to figure out how to make it work. Like in Fugazi, I used one guitar, one amp, one cord, and no pedals. The idea [being], what sounds can I get out of this instrument using only these few tools? How can I mimic an effect by using tone or volume or the way I finger things? I don't want more options. I want less options and more engagement.
>
> Back in 1979, I don't think I was particularly serious about having rules like "I'm not buying anything with keyboards on it," but [it] served as a good filter. An excellent way to avoid all of the yeses in the world when a simple "no" could suffice.
>
> \- Ian MacKaye [@mackaye2021]

## Punk

### Grateful Dead

> DJ: "You're not into any disco or New Wave or anything?"
>
> JG: "Of those things, of those choices, the New Wave\... the whole punk rock trip, I like it best. I like it best, I think. I like it because it has some energy, it has some spirit."
>
> JK: "There's the thing of, new music, people are tyring to play [crosstalk] there's thing, in San Francisco, 'cause that's where I was then, when everybody was just learning to play, there was a thing that was missing \--"
>
> JG: "The thing of people tryin' real hard.
>
> JK: "Tryin' to figure out how to make it work."
>
> JG: "Yeah, yeah. It's got the spirit, y'know, that's the thing. The nature of the spirit, if you had to describe the spirit, you could describe the spirit as being a little bit edgy, y'know, maybe sort of a violent spirit, y'know, sort of scaly and weird, maybe, but it's a spirit, and almost anything else lacks spirit of any sort." [@garcia1978]
>
> "It's always nice for the kids to have music. Punk music at it's finest was music that the kids could dig, and it was thoroughly obnoxious to adults. It had that great quality of being able to communicate to the kids and give them something to talk about, something to love, and also alienating everybody else. For me, it's the thing of, when you see young musicians, playing in a way, three chord rock 'n' roll, you know there's only one place for them to go, they're either going to get better, and finally learn how to tune their instruments, and do that kind of stuff, and evolve more, or else they're gonna shed more blood. For me, commitment counts, when you're on stage, and whether you're bleeding over every note, whether you're wounding yourself physically on whatever level, that's real."
>
> \- Jerry Garcia (1987) [@garcia1987]

### Fugazi

> My definition of punk is the free space. It's an area in which new ideas can be presented without having to go through the filtration or perversion of profiteering. So, if we're not worried about selling things, then we can actually think. The problem with new ideas is that they don't have audiences. And in terms of the marketplace, an audience equals clientele. If you have no audience, it's not profitable.
>
> But punk was an area, for me at least, where it didn't seem to matter. I didn't know any punk rocker who thought, 'I'm gonna make a living out of this.' The ones that did quickly left. What I received from the counterculture was a gift; the permission to create freely. And my reaction was to take care of this gift and keep it alive because it continues to give. Of course, there were some people who thought, 'Wow. If I polish it, I can sell it.' And then it ceases to be a gift.
>
> \- Ian MacKaye [@mackaye2015]

## Live Performances

Both Fugazi and the Grateful Dead would play different sets each show with no pre-defined setlist except for perhaps agreeing on which songs to start on.

### Grateful Dead

> 'We don't make up our sets beforehand. We'd rather work off the tops of our heads than off a piece of paper.'
>
> \- Jerry Garcia [@gans1985]

> 'Jerry Garcia and Bob Weir planned out only two things before each of the thousands of concerts they played together in the Grateful Dead: The song they would begin with and the song they would end with. Everything else was decided in the moment. "While Jerry was singing, I had plenty of time to think about the next tune that I was going to do," says Weir. "And while I was singing, he had plenty of time to think about what tune he was next going to do. We developed a feel for that."' [@doyle2020]

In a 1981 interview Jerry Garcia was asked what he thought about bands that played short live sets.

> "Fifteen-minute sets!" he marvels. "If I had to pay £8 for a 15-minute set I'd trip out\... The economics of it, I would feel so guilty. Even if I did a 45-minute show so packed with emotion and intensity and everything it needed to have I would still feel like, God it ain't fucking worth it. I don't want to burn anybody. People have to work to get their little money\... The best experiences I've had as an audience member is when I've seen a performer get excited and inspired and go over their time. Forget about time\...forget about time and then you can think hey! An hour and a half has gone by and it seems like ten minutes! That's the stuff!"
>
> \- Jerry Garcia [@garcia1981]

### Fugazi

> "We would get on stage with no setlist and then it was a freefall from there"
>
> "We worked out this system of hand signals... the one thing we knew was that Ian and I would trade vocals... usually it was like, if I started a song on guitar, then they would just have to be listening, if Brendan had to start it, I had to get him the message" 
>
> "The reason why it's a good idea and the reason why I'm shocked other bands don't do it is because it forces a level of attention on you and I kind of tightrope aspect to it that it doesn't let you relax in the set"
>
> \- Guy Picciotto [@picciotto2019]

## Lyrics

Both bands would often use lyrics that were open to interpretation and the songwriters were reluctant to explain the meaning of their songs.

### Grateful Dead

Most original Grateful Dead songs had lyrics written by full band members who did not perform on stage: Robert Hunter (104 songs) and John Perry Barlow (30 songs). Jerry Garcia wrote the lyrics to 5 songs, Bob Weir wrote the lyrics to 9 songs, Pigpen wrote the lyrics to 11 songs, Phil Lesh wrote the lyrics to 5 songs, Brent Mydland wrote the lyrics to 6 songs, Donna Godcheaux wrote the lyrics to 2 songs, and there are also 12 songs with lyrics credited to The Grateful Dead [@allan2024]

> In 'Long Strange Trip', there's a sequence in which Bar-Lev, after years of trying to get to Hunter, is granted an audience backstage after a solo gig. Hunter was always elusive and notoriously averse to discussing the meanings or sources of his compositions (though he was insistent about the songwriter's claim to royalties and credit, as the Internet undermined each). In the film, Hunter recites the lyrics to 'Dark Star' ("Dark star crashes, pouring its light into ashes / Reason tatters, the forces tear loose from the axis / Searchlight casting for faults in the clouds of delusion / Shall we go, you and I while we can / Through the transitive nightfall of diamonds?") and says, "What's unclear about that?" Then he kicks Bar-Lev out of his dressing room.
>
> "Meaning is often a subterfuge to distract the listener's attention from a writer's lack of multiple resources," Hunter writes. Then he presents an exegesis of his 1975 song 'Franklin's Tower', which became a great crowd-pleaser, in Garcia's hands. Hunter cites, among some of his references, Ben Franklin, the Constitution, Pete Seeger, the Bible, E. E. cummings, Bonnie Dobson, an Eastertide anthem called 'Roll Away the Stone', and the birth of his son.
>
> "Well, now that you know what I meant by it, it's no great shakes, is it?" Hunter concludes. "Mystery gone, the magician's trick told, the gluttony for 'meaning' temporarily satisfied, one can now take issue with my intent and avoid the song itself." [@paumgarten2019]

### Fugazi

Although all Fugazi songs are credited to Fugazi, the words to Fugazi songs were nearly always written by the person who would sing the song, mostly Ian MacKaye or Guy Picciotto with a few songs written and sung by Joe Lally.

> "I really do feel that my songs are not as inscrutable as they seem, but that is one I can't really defend, maybe it's the exception that proves the rule ... I think a lot of it has to do with engagement but also because I wrote the songs ... I always have this feeling that 'isn't that perfectly clear?' ... a song like let's say 'Last Chance for a Slow Dance' I remember reading something where someone was like 'this is just some more word-salad' ... that song is so personal and I felt so clear ... the part that goes like 'Flare fakes a flower / A burnt-out shower'. Someone was like, just thought that was gobbledigook or something, like to me it's the image of a flare being shot into the sky as a signal for help and it makes the form of a flower in the sky and it makes a burned-out shower of sparks falling and no-body can see it, it seems like such an obvious lonely image, and then people are like 'I really don't know what the fuck that means', it feels like a failure on one hand because it didn't work, but on the other hand I get kind of like 'shit, really? come on!' When I listen to a band I really like I enjoy that engagement, why would anyone go up to Bob Dylan and say like you know 'It's all Over Now Baby Blue' what's that song about? What is that song not about?"
>
> \- Guy Picciotto. [@mackaye2022]

## Politics

Both bands avoided linking themselves to specific political parties or candidates, with their political views being revealed by their lyrics, the ways they operated, and the causes they supported with benefit shows.

### Grateful Dead

> The Dead's attitude about politics would stay consistent, if unique. They probably did more benefits than any band ever, and frequently for explicitly political groups, but they'd sign nothing. They put their time where their beliefs were, but not their mouths. It was a policy designed to make "Grateful Dead" something separate, above life's daily shuffle. "It's our responsibility to keep ourselves free of those connotations," said Garcia years later. "I want the Grateful Dead experience to be one of those things that doesn't have a hook. We're all very antiauthoritarian. There's nothing that we believe so uniformly and so totally that we could use the Grateful Dead to advertise it." [@mcnally2003]
>
> Rock and politics, it was sort of inevitable but I never thought it was a good idea. In the Bay Area at that time there were two big philosophical pillars. There was the San Francisco approach, and there was the Berkeley approach. The Berkeley approach was politics, the endless argument, the pro-con kind of dualism. The San Francisco approach was psychedelic in the largest possible sense which is everything that happens happens, and that's the way it is. It lacked the polemic quality of the Berkeley discussion. For the Grateful Dead, from our point of view and the Haight-Ashbury point of view, the war and anti-war thing was all war. If you were talking pro-war or against the war you were talking war, your energy was going into the war, you were creating more war noise. For us the alternative was 'Hey, fuck the war. We don't need it, the world doesn't need it, nobody likes it, it never was much fun and maybe something else is what's called for. That's the broad philosophical difference. [@garcia1987a]

### Fugazi

Fugazi is strongly against violence, both at their shows and other types of violence such as war. They would stop playing if people started fighting at shows, and if necessary would give violent or abusive individuals a refund and escort them to the door personally.

> Fugazi has maintained their own unique brand of relativism, and non-dogmatic ideology, embodying the "principled anarchism" long adhered to by some of the best and generally forgotten musical outfits of recent history. They have been called elitist and humorless and yet have also been the recepients of an intense and broad fan loyalty unparalleled for such a strongly local entity. Most important, they embody the historically recognized anarchistic principles of individual and group autonomy, non-coercive relations, mutual aid, voluntary cooperation, and non-coercive advocacy. [@fairchild1995]

## Business Practices

Both the Grateful Dead and Fugazi are well-known exponents of particular ways of doing business as musicians.

### Grateful Dead

Whole books have been written about the way the Grateful Dead did business so what follows here is a very brief summary.

> "I thought it was interesting that they played 2,300-2,600 shows, and I was shocked. I thought they would have played around 5,000 shows, or something way beyond what they actually did."
>
> [interviewer] "Did you ever listen to them?"
>
> "No. But I think they're interesting structurally, in retrospect, for instance their business and the way they flexed their muscles in a good way. In the late 80s or early 90s, before Jerry died, they did these stadium shows, and they had their own ticketing system. They sold tickets directly, and at some point during this stadium tour, Ticketmaster didn't want them selling directly. They told The Dead that wasn't happening anymore, so the band told them, 'Okay, you're fired'. And then so Ticketmaster tried to backtrack. At least that's what I heard, I'm not sure if that's necessarily true. But anyways, they operated well, they weren't bullied by the machine. You get that sort of power by saying 'No'."
>
> \- Ian MacKaye (2013) [@mackaye2013]

> The Grateful Dead manner of writing songs is a very haphazard, hit-or-miss business. Nothing is nailed down. First they try their songs out in front of an audience. For most groups the song gets written and arranged, then it comes out on record and gets played on the radio. Only then does the band go out on the road and back up the record, basically lip-synching their own songs. The Dead, however, like to go out onstage and play a totally new song - something that they've just written or are still writing -long before it ever appears on an album.
>
> \- Rock Scully (former manager of the Grateful Dead) [@scully1996]

> The Dead were the first band to sign with a major label (Warner Brothers) with the condition that all publishing rights be kept within their own organization. Though the battles over this issue were furious, the band eventually prevailed. Unlike artists desperate to break into the business, the Grateful Dead did not care for recording an album if this would mean assigning away artistic control and copyright to their music. [@drobnik2000]

> Because the Dead had a tendency (especially in concert) to play extended jams in the middle of their songs, or to segue one song to the next via lengthy improvisational experimentation, they stood to receive lesser royalties for only recording two or three songs per album side. The answer to their problem came from the jazz world, where longer improvisational pieces are far more common. For years, jazz musicians' royalties were computed by the minute, not per song. The Dead took heed and convinced Warner Brothers to agree. [@drobnik2000]
>
> At the time (and still today), most artists would play for no longer than two hours, having written a performance limitation into their contracts with local promoters. But the Dead insisted that they be given at least four hours of stage time (many of their shows extended well beyond five hours in length). Furthermore, when major concert tour ticket prices began to skyrocket, the band refused to allow promoters to charge more than \$30 a ticket. [@drobnik2000]
>
> Whereas most bands hire independent crews for each tour, the Grateful Dead kept their employees on the payroll year-round. During the height of the band's success, crewmembers earned six figure salaries, were entitled to profit sharing, health benefit and retirement plans, and even were provided backstage daycare for their families. The loyalty that such treatment inspired undoubtedly resulted in high performance by their crew. [@drobnik2000]
>
> The Dead began this entire process of "own(ing) the marketing function themselves" in 1971, when the band inserted a notice in its Grateful Dead album, signed by Garcia and reading "DEAD FREAKS UNITE: Who are you? Where are you? How are you? Send us your name and address and we'll keep you informed." What followed were thousands of responses, leading later that year to the publication of a fan newsletter providing information on tour dates and requesting feedback about various quality concerns. In essence, the band adopted database marketing long before mainstream marketers took notice of this technique. [@drobnik2000]
>
> The band incorporated in 1973, the same year they launched Grateful Dead Productions and Round Records (which in itself was a radical decision for the time). Each band member became a CEO and obtained a seat on their board of directors, owning an equal share in profits and an equal vote in all decisions. Sam Hill and Glenn Rifkin comment that, "As collective CEOs of the business, [the Dead] owned the marketing function themselves. They never handed it off to a publicity firm or pushed it down into layers of a bureaucratic organization." In order to ensure that the band would continue to enjoy the benefits of artistic freedom and total control over its business direction and decisions, the Dead agreed that upon the death of any member, that person's shares would be reabsorbed by the organization and redistributed among the remaining CEOs. [@drobnik2000]
>
> Intellectual property law in the United States purports to maintain a balance between the interests of authors and those of the public. Necessarily, this objective implies that there is a barrier between the two. But since intellectual property is by its nature intangible property, the legal barriers defining its parameters are shifty at best. The Grateful Dead's philosophy was to eliminate many of these barriers, or at least to minimize their intrusiveness into the relationship the band shared with its fans. They accomplished this by first maintaining control over their legal rights in their property, and then by allowing near-complete access to it for those who wanted access. [@drobnik2000]
>
> ... audience members taped their shows from the start, and, until recently, copyright law has had little to say about bootleggers. But while other artists sought to sue bootleggers out of business under causes such as unfair competition, the Dead began to welcome them. Understanding that their bootleg albums were commanding a price on the market, the Dead opened their doors to every bootlegger who wanted to tape them, provided only that the tapes be bartered for other tapes. Sam Hill and Glenn Rifkin comment that "those with the most extensive tape collections became masters of their universe, and thus the open-taping decision fueled ticket sales." More than this, by flooding the market with their live product, the Dead eliminated the market in which for-profit bootleggers thrived. Furthermore, such an extension of trust magnified fan devotion, while the presence of countless bootlegs on the market created greater demand for touring. [@drobnik2000]
>
> The Grateful Dead's creed has always been, "When we are done with the music, [the fans] can have it." The extreme of this philosophy was the fact that the Dead often broadcast their concerts. Contrary to popular industry wisdom, the Dead customarily sent their live shows over empty airwaves. On any given night, audiences ranging in size from the people in the parking lot who could not get tickets to 20 million northern Europeans could hear what they were not able to see. [@drobnik2000]
>
> Though the Dead certainly have not overlooked their publicity interests, their stance is as relaxed and open as it is in regard to copyright and trademark control. Again, they function within the paradox of being in control of their affairs, and yet allowing outsiders to determine to a great degree the course of these. The band members have allowed their pictures to grace everything from Tshirts to ice cream. Aside from negotiating a licensing agreement and retaining rights to quality approval, they have allowed their fans to conduct the bulk of their publicity work. [@drobnik2000]

### Fugazi

Fugazi issued all of their recordings through the independent label Dischord Records which was founded by Ian MacKaye and Jeff Nelson in 1980 to put out records by Washington DC area bands. Dischord records are distributed either by mail order or by direct deals with local record shops, and are available anywhere in the world.

Fugazi have a corporation called Lunar Atrocities Limited.

> There's a Fugazi corporation which is called Lunar Atrocities Limited ... Protection from liability is the main reason to form a corporation, and for these guys it makes sense. If someone got hurt stage-diving and decided to sue, it would be a little harder to go after their personal assets." - Seth Martin [@brace1993].

## Summary

+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Topic                                              | Grateful Dead                                         | Fugazi                                |
+:===================================================+:======================================================+:======================================+
| Status                                             | Disbanded                                             | On indefinite hiatus                  |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Hometown                                           | San Francisco Bay Area, CA                            | Washington, DC                        |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Period of Activity                                 | 31 years (1965 - 1995)                                | 16 years (1987 - 2002)                |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Original songs played live                         | 151 [@amato2015]                                      | 94 [@mitrani2024]                     |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Cover versions played live                         | 190 [@amato2016]                                      | 1                                     |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Live performances                                  | 2325 [@gratefulseconds2016]                           | 1048 [@mitrani2024]                   |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Benefit shows                                      | 99 [@fatemusic2018]                                   | 80 [@klirs2018]                       |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Studio albums                                      | 13                                                    | 6                                     |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Songs on studio albums never performed live        | 2 [@benitez-eves2024]                                 | 0                                     |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Original songs performed live not on studio albums | 23 [@allan2021]                                       | 2 [@mitrani2024]                      |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Live albums                                        | 167 [@guinnessworldrecords2024]                       | 896 [@dischordrecords2024]            |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Live recordings                                    | 2065 [@relisten2024]                                  | 915 [@mitrani2023]                    |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Members (past and present)                         | 13                                                    | 4                                     |
|                                                    |                                                       |                                       |
|                                                    | [@thegratefuldead2023]                                | [@dischordrecords2024a]               |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Touring musicians                                  | 2                                                     | 1                                     |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Main instruments                                   | Vocals, Guitar, Keyboards, Bass, Drums.               | Vocals, Guitar, Bass, Drums           |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Other instruments                                  | Harmonica, Trumpet, Saxophone,                        | Trumpet, Clarinet, Soundboard effects |
|                                                    |                                                       |                                       |
|                                                    | Clavichord, Synthesizer, Computer                     |                                       |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| USA States played                                  | 45 [@golsen2022]                                      | 50 [@dischordrecords2024]             |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Countries played                                   | 12 [@twelvebarsoftwarellc2024]                        | 36 [@mitrani2024a]                    |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Venues played                                      | 592 [@twelvebarsoftwarellc2024]                       | 733 [@mitrani2024a]                   |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Average concert length                             | 3 hours [@teachrock.org2021]                          | 1 hour 17 minutes [@mitrani2024]      |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Length of longest concert                          | 4 hours 41 minutes [@internetarchivecontributors2014] | 2 hours and 3 minutes [@mitrani2024]  |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Average song length                                | 9 minutes [@teachrock.org2021]                        | 3 minutes 34 seconds [@mitrani2024a]  |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Longest live song rendition                        | 46 minutes 59 seconds [@twelvebarsoftwarellc2024]     | 13 minutes 52 seconds [@mitrani2024]  |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Average attendance                                 | 9604 [@gratefulseconds2016]                           | 938 [@mitrani2024]                    |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Largest live audience                              | 600,000 [@teachrock.org2021]                          | 15,000 [@mitrani2024]                 |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Total live audience                                | 22,330,524 [@gratefulseconds2016]                     | 983,881 [@mitrani2024]                |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+
| Funds raised for grass-roots organisations (USD)   | \$10,017,706 [@rexfoundation2024]                     | \$250,000 [@klirs2018a]               |
+----------------------------------------------------+-------------------------------------------------------+---------------------------------------+

## 

## References
