---
title: "in your memory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{in your memory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_in_your_memory.bib
link-citations: true
autosize: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

4 August 2025

## Introduction

Forensic Scene was played live for the first time at Colégio Leão XIII in Itaboraí, RJ, Brazil on 19 August 1994 [@fugazi1994], introduced by Guy with the comment "you're probably going to hate this song but we're going to play it anyway". Forensic Scene was one of 6 songs to debut on that 1994 tour of Brazil - including the full version of 'By You',  which had been played 5 times in 1993 as an instrumental. 

```{r echo=FALSE}
releases_data_input %>%
  filter(release=="red medicine") %>%
  select(song, date) %>%
  rename(debut = date) %>%
  arrange(debut)

```

The song got it's official release on Red Medicine which was released around the 12 May 1995 [@fugazi1995]. Fugazi played Forensic Scene at least 201 times with the last recorded version being on the 4 November 2002 in London, England - the final Fugazi show to date. It was the 5th most played song from Red Medicine. In live shows Forensic Scene tended to be used later in the set, it was often followed by Promises (31 times) or by Long Division (21 times).

There are several interviews where Guy Picciotto has commented on the song. In 1995 he only commented in outline:

> I don’t normally discuss what songs are actually specifically about. I feel like that the work that was put into creating the song is the expression and that people who listen to the song, if they’re interested in the song, create their own interpretations. I feel like, if it doesn’t make any sense, then I didn’t succeed when I wrote the song. I’m willing to accept that as a failure, but in general… Sometimes I’ll discuss a song like Cassavetes on In on the Kill Taker. I can say that that’s a song about the filmmaker John Cassavetes, because it seems to be fairly obvious, but Forensic scene… It’s like a less… It’s a more personal song. You can’t say that it’s about a specific person or something like that. It’s more about a situation. And I think it’s up to people to figure it out for themselves. [@mackaye1995]

In a 2025 interview with Brian Gathy for the End on End podcast he went into more detail:

> Brendan actually wrote the verse chords and I wrote the chorus chords... the verse chords seem simple but it's actually a very complicated thing. Brendan writes some very complicated guitar shit. His chordings are so wild and it's one of my favourite things that he ever wrote. The pleasure of learning to play that and then figuring out how to sing to it and then coming up with some big chorus ... it's one of my favourite things that Brendan and I worked on together, and it really came from those initial chordings that Brendan came up with, that are so interesting." [@picciotto2025]

> It was a very sad moment that produced that song, I don't really have anything more specific to say to it than that, ... it was kind of like that forensic analysis post-fact of the horrible thing that had just happened" [@picciotto2025]

The first time the song was recorded live, in August 1994, it was fully formed with both lyrics and music being more or less identical to the studio version that would be recorded in early 1995. This strongly suggests it had been written, rehearsed and possibly performed live some time earlier. The 1993 Japan / Australia / New Zealand Tour ended in Honolulu, Hawaii on the 14 December 1993 and the next show after that was Brendan's Going Away Party at Pirate House on 3 June 1994 (no recording exists). After that Brendan went to live in Seattle for several months, which means Forensic Scene was probably written sometime between December 1993 and May 1994.

## Title

The American Academy of Forensic Sciences has this to say about the word 'forensic':

> The word forensic comes from the Latin word forensis: public, to the forum or public discussion; argumentative, rhetorical, belonging to debate or discussion. A relevant, modern definition of forensic is: relating to, used in, or suitable to a court of law. Any science used for the purposes of the law is a forensic science. ... Forensic scientists may be involved anytime an objective, scientific analysis is needed to find the truth and to seek justice in a legal proceeding. [@aafs2022]

'Scene' has several possible meanings [@merriam-webster2025]:

-   one of the subdivisions of a play

-   a stage setting, or a real / imaginary view suggesting a stage setting

-   the place of an occurrence or action

-   an exhibition of anger or indecorous behavior

-   a sphere of activity or situation

## Lyrics

The lyrics from the studio recording will be referred to here [@fugazi1995] - there are some variations in live recordings. The lyric sheet uses all lower case with no line breaks, here I'm going to use lower case and line breaks based on how it is sung.

Notice how many of the lines are couplets and some of them rhyme although the rhymes seem forced at times: gentle / sentimental; without / now; would / wound; lay / anyway; unclean / scene; instant / distance.

### in your memory not so gentle

"in your memory, not so gentle": something not so gentle in memory of someone.

### perversify slash sentimental

If 'versify' means to turn something into verse, then 'perversify' would mean to make something perverse.

'perversify slash sentimental' suggests being by turns perverse and sentimental about something.

### caught between and caught without

Unexpectedly in a difficult or uncomfortable situation.

Words are omitted leaving the listener to fill in the blanks.

Several sayings come to mind:

-   caught between a rock and a hard place

-   caught without an umbrella.

There is a song called 'Caught without an umbrella' by Spearhead (Michael Franti's band) from 1994 that is about a failed suicide attempt [@spearhead1994].

### i'm a failure not your failure now

This is on me, it's not your responsibility anymore. Taking responsibility for a bad situation or for having done something wrong. Expresses humility and perhaps also regret.

'Not your failure now' hints at a family or intergenerational aspect to the situation, for instance a person with a troubled childhood whose parents might have felt that they failed in giving them a good upbringing. It also suggests finality, the end of a relationship.

### in and out just like a knife would

Something that is not a knife that penetrates like a knife - maybe a needle, an insect sting or bite. This could refer to drug use.

### tax the flesh and leave a cheap wound

Unusual usage of tax. The only similar usage I could find was in an academic paper on the etymology of the word 'dismal', which suggests a possible connection to some religious practices such as fasting and flagellation:

> in our sentence, desmar las carns, the verb desmar is used to refer, not to time, but to the flesh... 'to decimate, to diminish'. "now is the time to 'tax' the flesh." The phrase decimare carnem is but a variant of the phrase exuere carnem which, according to Saint Augustine, expresses the duty of the christian who wants to follow Christ. [@spitzer1942]

'Cheap wound' implies that it is something small, not a serious injury, which seems consistent with the idea of a needle being used for injections.

### fucked and fucking love and lay

This brings to mind press coverage of celebrities, in particular their misfortunes and relationships. For instance, from the 1992 Vanity Fair article 'Strange Love':

> Are Courtney Love, lead diva of the postpunk band Hole, and her husband, Nirvana heartthrob Kurt Cobain, the grunge John and Yoko? Or the next Sid and Nancy? [@hirschberg1992]

### congenitally fractured anyway

Congenital in this context would mean 'existing at or dating from birth', 'an essential characteristic', or 'being such by nature' [@merriam-webster2025a].

A continuation of the previous line, alluding to commentary and speculation about other people's lives and what made them do what they did.

### unclipped unclean

Unclipped nails, dirt or blood under nails could be part of a forensic investigation in order to establish responsibilites.

### this forensic scene's

Forensic seems to be used here in the sense of people trying to figure out what happened, why it happened, and who is to blame.

Most if not all of the meanings of 'scene' seem relevant here. Scene as in a place where something happened, a community of bands and musicians, and an actual or imagined visual perspective.

### all played out

'played out' means exhausted, spent, finished. 'played out' also resonates with one of the meanings of scene - a scene of a play, perhaps alluding to something dramatic about the subject of the song.

### the defense rests

In a trial the defense rests when they have no more evidence to present at that stage. This implies that the process is moving towards some sort of closure.

### and sorry's just a no shit sherlock mouth talk con job

Alludes to a situation that is so serious that an apology is inadequate and perhaps inappropriate

'no shit sherlock', meaning something that is obvious and does not need to be said, ironically references Sherlock Holmes, Arthur Conan Doyle's famous detective character who made use of deductive reasoning to infer the truth behind the observed facts of his cases.

'mouth talk con job' indicates insincerity or deception.

### from your memory to this instant

From thinking about someone in the past to the present moment.

### worked so hard for working distance

This alludes to being too close to something, and perhaps trying to get more distance in order to get a better perspective.

'working distance' in photography refers to the distance between the lens and the subject being photographed [@carnathan2014]. The greater the working distance, the easier things are in terms of lighting and composition. If the distance between the lens and the subject is shorter than the working distance, the subject will be out of focus.

### like a mouth too late to shut

This is a metaphor for having done or said something wrong, implies some degree of regret.

### i'm a failure not your failure now

Outside of the chorus, this is the only line that is repeats, which gives it emphasis.

## Live variations

The 1995 show at Queen's Hall, Belfast contains a grungy extended ending with Ian and Guy singing "over and over and over and over and over and over and over and over and over and over..." [@fugazi1995a]

Many live performances of Forensic Scene include an additional line just before the instrumental break, some of these are easier to make out than others, for instance "take a walk, sister" [@fugazi1996], [@fugazi1996a].

## Press Cuttings

What happened between December 1993 and May 1994 that could be described as a 'horrible thing' and might have prompted the composition of this song? One event in particular comes to mind.

> NIRVANA SINGER FOUND DEAD By Richard Harrington and Richard Lei. April 9, 1994. Kurt Cobain, who led the grunge band Nirvana to superstardom yet reviled his own success, was found shot in the head in his Seattle home yesterday, an apparent suicide. [@harrington1994]

> Love reconstructed the rest of the tragedy for MTV: Cobain drew a chair up to a window overlooking the Puget Sound, sat down, took some more drugs (most likely heroin), pressed the barrel of the 20-gauge shotgun to his head and — evidently using his thumb — pulled the trigger. [@strauss1994]

The rest of the press cuttings in this section have been chosen because they seem relevant, and some of them are potentially related to the lyrics of Forensic Scene. They are presented in approximately chronological order.

> With his untimely death, Cobain enters a sad rock pantheon that includes Jim Morrison and Janis Joplin, along with Seattle musicians Jimi Hendrix, Stephanie Sargent of the band 7 Year Bitch, and Anthony Wood of Mother Love Bone. "Now he's gone and joined that stupid club," Cobain's mother, Wendy O'Connor, told reporters yesterday. "I told him not to join that stupid club." [@harrington1994]

> All three members of Nirvana came from broken homes. Cobain, son of an auto mechanic and secretary who divorced when he was 8, did not speak to his father for eight years -- until signing his breakthrough contract with Geffen in 1990. As a shy teenager, he became enthralled with punk rock and began collecting records and playing guitar, though he saw himself as a rhythm guitarist -- out of the limelight. When Cobain was still a teenager, his father had forced him to pawn his guitar and join the Navy; instead Cobain reclaimed the guitar and left home. [@harrington1994]

> On behalf of Dave, Pat, and I, I would like to thank you all for your concern at this time. We remember Kurt for what he was: caring, generous, and sweet. Let's keep the music with us. We'll always have it... Forever. Kurt had an ethic towards his fans that was rooted in the punk rock way of thinking. No band is special, no player royalty. But if you've got a guitar and a lot of soul just bang something out and mean it. You're the superstar. Plugged in the tones and rhythms that are uniquely and universally human: music. Heck... use your guitar as a drum, just catch the groove and let it flow out of your heart. That's the level Kurt spoke to us on: in our hearts, and that's where he, and the music, will always be, forever. [@novoselic1994]

The following extracts are from Courtney Love's pre-recorded eulogy which was played at a public memorial service on 10 April 1994 at the Seattle Center. In the recording she intersperses parts of Kurt Cobain's suicide note with her own comments.

> *This note should be pretty easy to understand. All the warnings from the punk rock 101 courses over the years since my first introduction to the, shall we say, ethics involved with independence and embracement of your community, has proven to be very true. I haven't felt the excitement of listening to, as well as creating music, along with really writing something, for too many years now. I feel guilty beyond words about these things - for example, when we're backstage and the lights go out and the roar of the crowd begins, it doesn't affect me the way in which it did for Freddie Mercury, who seemed to love and relish the love and adoration of the crowd.* [@cobain1994]
>
> Well, Kurt, so fucking what - then don't be a rock star you asshole. [@love1994]
>
> *Which is something I totally admire and envy. The fact, I can't fool you, any one of you, it simply isn't fair to you or to me. The worst crime I could think of would be to put people off by faking it, pretending as if I'm having 100% fun.* [@cobain1994]
>
> No Kurt, the worst crime I can think of is for you to just continue being a rock star when you fucking hate it, just fucking stop. [@love1994]

> And I'm laying in our bed, and I'm really sorry. And I feel the same way you do. I'm really sorry you guys. I don't know what I could have done. I wish I'd been here. I wish I hadn't listened to other people, but I did. [@love1994]

On the aftermath, including comments from counselors and help-line workers concerned about young people at risk of suicide.

> The days after Cobain’s death were filled with grief, confusion and finger pointing for all concerned. “Everyone who feels guilty, raise your hand,” Love told MTV the morning after Cobain was found. She said she was wearing Cobain’s jeans and socks and carrying a lock of her husband’s blond hair. According to Billig, a doctor was summoned to stay with Love at all times. [@strauss1994]

> "Cobain's suicide does send a message, certainly, particularly for youth on the edge. These musicians are kids' role models. They find their identity in the music. It's an angry-tone music, angst filled, hopeless to a certain degree. But Nirvana's not suggesting everybody go out and shoot themselves. The main thing is to focus on what's causing suicidal feelings, what the issues are. Suicide is a permanent solution to a temporary problem. The intensity is real subjective, the feeling that it will never change. But this is what not to do. If you're feeling overwhelmed, use the phone as a tool. Reach out and get help." [@anders1994]

> "When we see this popular person, we tend to think that they have everything. And yet something about that still doesn't make their life okay -- so the message is that nothing makes you happy enough to make you want to live. Young people may think, 'He did it. That means that I could also do it.' This reminds me of when Magic Johnson announced he was HIV-positive: It creates more intensity of vulnerability. Like, 'My God, it can also happen to me.' [@anders1994]

> "I think there's a lot of anger in a suicide, generally. And I always think that the anger is turned against the self. For kids who are so turned on by him, maybe it would help them to know that they are not the object of his anger -- probably his wife or his mother or father are."
>
> Gene Gordon, professor of psychiatry, George Washington University Medical School [@anders1994]

The following extracts are about works of music or art that were produced later.

> It has been 2 years, 2 months and 15 days since Cobain, the leader of Nirvana, was found dead in a small room above his garage, and songs influenced by his suicide continue to stream out of songwriters' pens from Seattle to Dublin. There are scores of songs known to be about Kurt Cobain – among them ones by Neil Young, Pearl Jam, the Cranberries, R.E.M., Julian Cope, Vernon Reid and the Tragically Hip -- and dozens more believed to be about Cobain, though the bands aren't telling (the Foo Fighters, Filter). [@strauss1996]

There is a painting made in 2004 called 'Untitled (forensic scene)' by Jordan Kantor [@kantor2004], an artist whose "painting practice is fueled by breakdowns or gaps in perception, often caused by the intermediary presence of the camera". The painting is part of a series where "he has relied on the spotty photographic record of the scene of Kurt Cobain's suicide to piece together impressions of the place and event, drawn by the intense public interest in it as well as the compelling privacy and lack of access that discouraged further knowledge."

> *Untitled (forensic scene)* (2004), is a painting of a zoomed-in view of what you might see if the policeman were to move aside. It’s the only painting in the cluster, and, on the one hand, a fragmentary recasting of Manet’s *The Dead Toreador*, on the other hand, a commentary on how the artist sees his subject. I like to think the kneeling figure is Kantor himself, winking off camera at the viewer: *This is what we do*. [@rader2017]

## Discussion

Guy's comment that "it’s a more personal song" [@mackaye1995] is helpful for understanding the lyrics. With this in mind I understand "in your memory" as about the memory of someone, it's not about somebody else remembering something. Furthermore, "from your memory to this instant" suggests that the person only exists in memories - they are dead.

Guy noted that "You can’t say that it’s about a specific person or something like that. It’s more about a situation." [@mackaye1995]. However, considering the time interval in which Forensic Scene would have been written, and Guy's comment that "it was a very sad moment that produced that song ... it was kind of like that forensic analysis post-fact of the horrible thing that had just happened" [@picciotto2025], it seems likely that the lyrics to Forensic Scene were written about the aftermath of Kurt Cobain's suicide. Fugazi were certainly aware of Nirvana and probably had met Kurt Cobain in person on a number of occasions, as well as being part of an extended scene of bands and musicians. They would have had shared interests and friends in common, so they would probably have been affected by his death. 

> If baby boomer cultural dominance was like a dam holding back the generation after it, that dam finally burst the day of Nevermind’s release in September of 1991. At the time, Fugazi was on tour in Australia promoting Steady Diet of Nothing. “It was like our record could have been a hobo pissing in the forest for the amount of impact it had,” says Guy Picciotto. “Nevermind was so huge, and people were so fucking blown away. We were just like, ‘What the fuck is going on here?’ It was so crazy. On one hand, the shows were bigger, but on the other hand, it felt like we were playing ukuleles all of a sudden because of the disparity of the impact of what they did.” [@azerrad2012]

The lyrics are cryptic enough to make the song flexible and open to other interpretations as well. Several discussions of the song have characterized it as being about the break-up of a relationship, for instance [@wright] and [@gathy]. In 1995 at Vassar College Walker Fieldhouse, Poughkeepsie, NY, Guy gave the following introduction:

> Did you ever have a couple of days where like everything you do is some incredibly stupid-ass mistake? Well, I've had a couple of days like that and this is a song for me, called Forensic Scene. [@fugazi1995a]

## References
