---
title: "92 SONGS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Livehistory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params: 
  output_dir: "../docs"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
output_dir <- params$output_dir
```

```{r, include = FALSE}
library(Repeatr)
```

## A visualisation of Fugazi history

This page offers a visualization of the songs that Fugazi performed live using metadata from the [Fugazi Live Series](https://www.dischord.com/fugazi_live_series).  

The visualization uses results of a previous analysis process that can be found in the summary data of the  [Repeatr](https://github.com/alexmitrani/Repeatr) package.  

Let's get the data, limit it to the columns that we will be using, and have a look at the first few rows. 

```{r }

mygraphdata <- summary %>%
  select(song, launchdate, chosen, release, rating)

head(mygraphdata)

``` 

Now let's graph the data. The idea is to plot each song as a bubble, with the x-coordinate given by the launch date, and the y-coordinate given by the rating calculated from the choice modelling of the Fugazi Live Series data. The size of the bubble will be proportional to the number of times the song was played live, while the color of the bubble will indicate the associated release in the band's discography.  There will be a lot of information packed into this graph!    

```{r }

p <- mygraphdata %>%
  ggplot( aes(x=launchdate, y=rating, size = chosen, color=release, label=song)) +
  geom_point(shape = 1) +
  theme_bw()

ggplotly(p)


```   

The graph is interactive: 

- hover over a song to see specific details about the song

- tap on a release in the legend to hide the corresponding
bubbles, tap on the release again to reveal them once more.  This is useful to focus on one or more specific releases by hiding the others.   


Enjoy!


