---
title: "Link Tracks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Link Tracks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
```

## Introduction

A series of brief explorations of the Fugazi Live Series data.  

## What was Fugazi's biggest show?  

```{r}
attendancedata <- othervariables %>%
  filter(is.na(attendance)==FALSE) %>%
  mutate(attendance = as.integer(attendance)) %>%
  mutate(date = as.Date(date, "%d-%m-%Y")) %>%
  mutate(year = lubridate::year(date)) %>%
  select(year, date, venue, attendance)

maxattendance <- max(attendancedata$attendance)

maxattendance

attendancedata %>%
  filter(attendance == maxattendance)
```
The biggest show was the [Food Not Bombs 20th Anniversary](https://www.dischord.com/fugazi_live_series/san-francisco-ca-usa-60400) on the 4th of June 2000 at Mission Dolores Park in San Francisco, with an estimated attendance of `r maxattendance` people.  It seems that the show was recorded but it is not available yet as part of the Fugazi Live Series.  [There is a video of the show on youtube](https://www.youtube.com/watch?v=0qYvHzS7v-Q).

## What was Fugazi's longest tour?  

```{r}

meanattendance <- othervariables %>%
  filter(is.na(tour)==FALSE) %>%
  mutate(attendance = ifelse(is.na(attendance)==TRUE, 100, attendance)) %>%
  group_by(year) %>%
  summarise(meanattendance = mean(attendance)) %>%
  ungroup()

toursdata <- othervariables %>%
  filter(is.na(tour)==FALSE) %>%
  left_join(meanattendance) %>%
  mutate(attendance = ifelse(is.na(attendance)==TRUE,meanattendance,attendance)) %>%
  group_by(tour) %>%
  filter(is.na(date)==FALSE) %>%
  summarise(start = min(date), end = max(date), shows = n(), duration = as.numeric((end - start)), attendance=sum(attendance)) %>%
  ungroup() %>%
  arrange(desc(shows))

  toursdata <- toursdata %>%
    mutate(meanattendance = as.integer(attendance / shows)) %>%
    arrange(start)

  toursdata <- toursdata %>%
    mutate(start = as.Date(start, "%d-%m-%Y")) %>%
    mutate(end = as.Date(end, "%d-%m-%Y"))

  toursdata$startyear <- lubridate::year(toursdata$start)
  toursdata$endyear <- lubridate::year(toursdata$end)

  toursdata$duration <- as.integer(toursdata$duration)
  toursdata$attendance <- as.integer(toursdata$attendance)
  
  toursdata <- toursdata %>%
    arrange(desc(shows))

head(toursdata, n=10)

longest_tour <- toursdata %>% arrange(desc(shows)) %>% slice(1)
biggest_tour <- toursdata %>% arrange(desc(attendance)) %>% slice(1)

```

On the `r longest_tour$tour`, between `r format(longest_tour$start)` and `r format(longest_tour$end)`, Fugazi played `r longest_tour$shows` shows over `r longest_tour$duration` days, with a total attendance of `r format(longest_tour$attendance, big.mark=",")` people. This tour wasn't the longest in terms of the number of days, or the biggest in terms of total attendance (the `r biggest_tour$tour` had a total attendance of `r format(biggest_tour$attendance, big.mark=",")` people), but it was the longest tour in terms of the number of shows.

## Leads and lags

Most Fugazi songs were performed live for some time before being released on an album or EP. These lead times were often very considerable, measured in months or years. Were there any exceptions, songs whose live launch dates lagged behind the corresponding release dates? To find out, let's start by getting the data on the releases and the corresponding release dates. 

```{r}
releasedates <- releasesdatalookup %>%
  select(rid, release_title, release_date)

mydf <- songvarslookup %>%
  left_join(releasedates) %>%
  left_join(songidlookup)
mydf <- mydf %>%
  select(songid, title, rid, release_title, release_date) %>%
  arrange(songid)
head(mydf)

```
Now let's calculate leads and lags by getting summary data on the songs and comparing the song launch dates to the corresponding release dates.

```{r}
mysummary <- Repeatr::summary %>%
  left_join(mydf) %>%
  mutate(lead = release_date - launchdate) %>%
  select(title, launchdate, release_title, release_date, lead) %>%
  arrange(lead)

head(mysummary, n = 10)

lagged_songs <- mysummary %>%
  filter(as.numeric(lead) < 0) %>%
  mutate(lag_days = abs(as.numeric(lead)),
         title = tools::toTitleCase(title),
         release_title = tools::toTitleCase(release_title))

n_lagged <- nrow(lagged_songs)

lagged_songs_text <- paste(
  paste0(lagged_songs$title, " which was first played live ", lagged_songs$lag_days,
         " day", ifelse(lagged_songs$lag_days == 1, "", "s"), " after the launch of ", lagged_songs$release_title),
  collapse = ", and "
)
```

Surprisingly, there `r ifelse(n_lagged==1, "seems to be only one song", paste0("seem to be only ", n_lagged, " songs"))` whose live `r ifelse(n_lagged==1, "debut lagged", "debuts lagged")` behind the corresponding release date`r ifelse(n_lagged==1, "", "s")`: `r lagged_songs_text`. What was the average lead time for all Fugazi songs with a corresponding release?

```{r}
mean_lead <- mean(mysummary$lead)
mean_lead
```
That is over `r floor(as.numeric(mean_lead)/365)` years, but perhaps the mean is biased upwards by a few extreme values...

```{r}
mysummary <- mysummary %>%
  select(title, launchdate, release_date, lead) %>%
  arrange(desc(lead))

head(mysummary, n = 10)
```
The median lead time is probably a more reliable indicator for how long Fugazi would tend to play a song live before it featuring on a discographical release.  

```{r}
median_lead <- median(mysummary$lead)
median_lead
```
We have discovered several interesting things:

1) `r ifelse(n_lagged==0, "No Fugazi song's live debut lagged behind its corresponding release.", paste0(ifelse(n_lagged==1, "The only Fugazi song whose live debut lagged behind its corresponding release was ", "The Fugazi songs whose live debuts lagged behind their corresponding releases were "), lagged_songs_text, "."))`

2) The median lead time for the live performance of a Fugazi song ahead of its corresponding discographical release date was approximately `r round(as.numeric(median_lead)/365, 1)` year`r ifelse(round(as.numeric(median_lead)/365, 1)==1, "", "s")`: `r as.integer(median_lead)` days.

## At which venues did Fugazi play the most?  

Listening to the Fugazi Live series in chronological order, the band returns to some venues again and again over the years. Let's have a look at the venues with the largest numbers of Fugazi shows.    

```{r}
venuesdata <- othervariables %>%
  mutate(year = year(date)) %>%
  mutate(city = ifelse(flsid=="FLS1053", "Bremen", city)) %>%
  mutate(country = ifelse(flsid=="FLS1053", "Germany", country)) %>%
  filter(is.na(venue)==FALSE & is.na(city)==FALSE & is.na(country)==FALSE) %>%
  group_by(venue, city, country) %>%
  summarize(shows = n(), from=min(year), to = max(year)) %>%
  select(venue, city, country, shows, from, to) %>%
  arrange(desc(shows)) %>%
  ungroup()

head(venuesdata, n = 10)

venues_over_10 <- venuesdata %>%
  filter(shows > 10) %>%
  arrange(desc(shows))

top_venue_years <- shows_data %>%
  filter(venue == venues_over_10$venue[1]) %>%
  distinct(year) %>%
  arrange(year) %>%
  pull(year)

top_venue_missing_years <- setdiff(min(top_venue_years):max(top_venue_years), top_venue_years)

```
The top 10 venues are all in the USA. `r venues_over_10$venue[1]`, in `r venues_over_10$city[1]`, is `r ifelse(nrow(venues_over_10)==1, "the only venue", paste0("the top venue, one of ", nrow(venues_over_10), " venues"))` with more than 10 shows. Fugazi played shows there `r venues_over_10$shows[1]` times between `r venues_over_10$from[1]` and `r venues_over_10$to[1]`, only missing `r length(top_venue_missing_years)` year`r ifelse(length(top_venue_missing_years)==1, "", "s")` (`r paste(top_venue_missing_years, collapse=", ")`).

Let's have a look at the top 10 overseas venues.  

```{r}
overseas_venuesdata <- venuesdata %>%
  filter(country!="USA" & shows>=4) %>%
  arrange(desc(shows))

head(overseas_venuesdata, n = 20)

top_overseas_shows <- max(overseas_venuesdata$shows)
top_overseas <- overseas_venuesdata %>% filter(shows == top_overseas_shows)
other_overseas <- overseas_venuesdata %>% filter(shows < top_overseas_shows)

countries_with_article <- c("Netherlands", "Czech Republic")
top_overseas_country_text <- ifelse(top_overseas$country %in% countries_with_article,
                                     paste("the", top_overseas$country), top_overseas$country)

```

Overseas, the `r ifelse(nrow(top_overseas)==1, "venue", "venues")` with most Fugazi shows `r ifelse(nrow(top_overseas)==1, "was", "were")` `r paste(paste0(top_overseas$venue, " in ", top_overseas_country_text), collapse=" and ")`, `r ifelse(nrow(top_overseas)==1, "with", "each with")` `r top_overseas_shows` shows. There were `r nrow(other_overseas)` other overseas venues with `r other_overseas$shows[1]` shows. Proud to see that the number 1 Fugazi venue in the UK was the Newcastle Riverside, in my home town, which was where I saw them play in 1990!

Let's have a quick look at the frequency distribution of the number of shows at each venue.  

```{r}
number_venues <- nrow(venuesdata)

cat(paste0("\n \n There are ", number_venues, " venues in the Fugazi Live Series data. \n \n"))

overview_venuesdata <- venuesdata %>%
  group_by(shows) %>%
  summarize(venues = n()) %>%
  mutate(percentage = round(100*venues/number_venues, digits = 3)) %>%
  arrange(desc(shows)) %>%
  ungroup()

head(overview_venuesdata, n = 11)

pct_shows <- function(n) {
  v <- overview_venuesdata$percentage[overview_venuesdata$shows == n]
  if (length(v) == 0) 0 else v
}

pct_5_or_more <- sum(overview_venuesdata$percentage[overview_venuesdata$shows >= 5])

```

Fugazi played at `r number_venues` venues but played at `r pct_shows(1)`% of them only once, twice at `r pct_shows(2)`% of venues, 3 shows at `r pct_shows(3)`% of venues, and 4 shows at `r pct_shows(4)`% of venues. Only `r pct_5_or_more`% of venues had 5 or more shows.

## In which city did Fugazi play at the most venues?

```{r}
mydf <- othervariables

venues <- mydf %>% 
  group_by(city, venue) %>% 
  summarize(shows = n()) %>% 
  ungroup()

venues_per_city <- venues %>% 
  group_by(city) %>% 
  summarize(venues = n()) %>% 
  arrange(desc(venues)) %>% 
  ungroup()

venues_per_city

```
The city where Fugazi played at the most venues was `r venues_per_city$city[1]`, followed by `r venues_per_city$city[2]`, and `r venues_per_city$city[3]`.

## Did Fugazi pick songs to perform randomly?

If songs were picked at random from all available songs, the performance count for each song should build up in a similar way over time.  This can be checked by plotting the cumulative performance counts for a selection of songs.  First, let's get the data into a suitable format, so that it will be easy to count the number of times each song was played. 

```{r}

mydf <- Repeatr1 %>% select(date, title)

mydf <- mydf %>%
  group_by(date, title) %>%
  summarize(count=n()) %>% ungroup()

mydf_wide <- mydf %>%
  pivot_wider(names_from = title, values_from = count, values_fill = 0)

head(mydf_wide)

```

Next, let's transform the variable for each song into a cumulative count of the number of times the song was played.  

```{r}

mydf_wide2 <- mydf_wide

for(colindex in 2:ncol(mydf_wide2)) {
  
  mydf_wide2[,colindex] <- cumsum(mydf_wide2[,colindex])     
  
}

head(mydf_wide2)

```

Now, we can reformat the data again to make it a long list of song counts that show how the number of times each song was performed built up over time.  

```{r}

mydf_long <- mydf_wide2 %>%
  pivot_longer(!date, names_to = "title", values_to = "count") %>%
  filter(count>0)

head(mydf_long)

```

Now we are in a position to graph the data so see how the performance counts evolved over time.  

Here is a selection of interesting songs to look at.  It would be too cluttered to plot all the songs at once!

```{r}

mydf_long %>%
  filter(title=="furniture" | title=="waiting room" | title=="shut the door" | title=="kyeo" | title=="polish" | title=="steady diet" | title=="smallpox champion" | title=="birthday pony" | title=="break" | title=="nightshop") %>%
  ggplot(aes(date, count, color = title)) +
  geom_line()

```

Another approach is to group the songs by release.

```{r}
releases_lookup <- Repeatr1 %>%
  group_by(title, release_title) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  select(title, release_title)

mydf_long <- mydf_long %>%
  left_join(releases_lookup)

mydf_long %>%
  filter(release_title=="steady diet of nothing") %>%
  ggplot(aes(date, count, color = title)) +
  geom_line()

cumulative_song_counts <- mydf_long %>%
  select(date, title, release_title, count)
```


Returning to the initial question of whether Fugazi picked songs randomly, the short answer is no, it doesn't look like it. Their choices indicate preferences that became clearer over time, as can be seen in the above graphs. They sometimes stopped playing certain songs for prolonged periods of time. For instance, there were lengthy periods where KYEO and Furniture were not being played at all, and then they were brought back and played again. Other songs such as Polish were dropped and did not return. Some of the decisions of what songs to play were probably spontaneous but spontaneity is not necessarily the same as randomness, although there might be some randomness involved. Fugazi tried [a different system for picking songs](https://www.dischord.com/fugazi_live_series/waukesha-wi-usa-61590) once but it only lasted about one song!    

There is an interactive version of the above graph available [here](https://alexmitrani.shinyapps.io/Repeatr-app/) - have fun!


## Mapping the Fugazi Live Series

I have tried my best to locate all the shows on an interactive, online map. I started with the information on the [Fugazi Live Series](https://www.dischord.com/fugazi_live_series) website, including any flyers and comments. Google Maps would find some places quickly but others would be more difficult - many of the venues are no longer there. The search was broadened using sites like [setlist.fm](https://www.setlist.fm/), which sometimes have addresses for old venues, reading online publications, and reaching out to people who might remember where the venues were located.  

To the best of my knowledge, all of the shows have been located to a reasonable degree of accuracy. The results can be found [here](https://alexmitrani.shinyapps.io/Repeatr-app/).

If you see an error that needs correcting, please open an issue on the [Repeatr GitHub](https://github.com/alexmitrani/Repeatr). 

## Sound Quality

Here is a summary of the sound quality ratings of shows in the Fugazi Live Series. A few of the shows are on [archive.org](https://archive.org/details/Fugazi) but not on the main site at this time. The Fort Reno shows include recordings for every level on the scale from Poor to Excellent. 

```{r}
sound_quality_ratings <- shows_data %>% 
  filter(is.na(sound_quality)==FALSE) %>% 
  group_by(sound_quality) %>% 
  summarize(shows = n()) %>% 
  ungroup()

totalshows <- sum(sound_quality_ratings$shows)

totalrow <- as.data.frame(t(c("Total", totalshows)))

names(totalrow) <- c("sound_quality", "shows")

sound_quality_ratings <- as.data.frame(rbind(sound_quality_ratings, totalrow))

sound_quality_ratings <- sound_quality_ratings %>%
  mutate(shows = as.numeric(shows)) %>%
  mutate(index = case_when(
    sound_quality == "Excellent" ~ 1,
    sound_quality == "Very Good" ~ 2,
    sound_quality == "Good" ~ 3,
    sound_quality == "Poor" ~ 4,
    sound_quality == "Total" ~ 5
  )) %>%
  arrange(index) %>%
  relocate(index)

sound_quality_ratings <- sound_quality_ratings %>%
  mutate(percentage = round(100*(shows / totalshows),1))

sound_quality_ratings



```

## Three Repeats, but Only One Two for Tuesdays

There is a small number of shows where the same song was performed twice, for different reasons. Some are easier to find than others and there are a few false positives. Let's have a look, shall we? 

```{r}
two_for_tuesdays <- Repeatr1 %>%
  filter(tracktype==1) %>%
  group_by(date, gid, title) %>%
  summarize(count = n()) %>% 
  ungroup() %>% 
  filter(count>=2)

two_for_tuesdays

```
'Break In' was played twice [in Annapolis in 1988](https://www.dischord.com/fugazi_live_series/annapolis-md-usa-20688) for musical reasons - Ian broke a string the first time through. Neither 'Reclamation' nor 'Bed for the Scraping' are true duplicates - in both cases the song was so badly interrupted the interruption was treated as a separate track splitting the song in two MP3 tracks. 'Great Cop' was actually played twice due to a conflict with security staff at [the 1998 gig in Richmond, Virginia](https://www.dischord.com/fugazi_live_series/richmond-va-usa-51198). The 1998 Sanctuary Theater show in Washington DC features two versions of both 'Closed Captioned' and 'Foreman's Dog' but this is because two songs from the soundcheck are included in the MP3 download as a bonus.     

The real "Two for Tuesdays" was at [a 1991 gig in Birmingham, Alabama](https://www.dischord.com/fugazi_live_series/birmingham-al-usa-52191), where the show notes comment "Featuring the one-time attempt of our 'Two for Tuesday' gag. No one appeared to notice, so we shelved the idea." On that occasion, the song "Greed" was played twice, and the gag was announced immediately after the double performance: "Two for Tuesday, Two for Tuesday". Perhaps this was a little too subtle? 'Greed' is such a short song you can fit two renditions into the time it would take to play a normal song. I was specifically looking for the shows where the same song was played twice and this one didn't come up on the above list because 'Greed' followed by 'Greed' was treated as one song in the track listing and one MP3 track - perhaps because the person who mastered the recording didn't notice either, or perhaps they wanted the gag to live on in the digital realm? I also missed it the first time I listened to the show. I bet more people would have noticed if it had been 'Suggestion'! 


Thanks.





