---
title: "Nice New Outfit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nice New Outfit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_nice_new_outfit.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

11 September 2023

## Introduction

Nice New Outfit is the third track on Fugazi's Steady Diet of Nothing album which was recorded at Inner Ear studios in January 1991 [@fugazi1991].

Judging from the dates of debut live performances in the Fugazi Live Series, the songs on Steady Diet of Nothing were written at different times in a period of about 3 years and 3 months, from October 1987 to January 1991.

```{r echo=FALSE}
releases_data_input %>%
  filter(release=="steady diet of nothing") %>%
  select(song, date) %>%
  rename(debut = date) %>%
  arrange(debut)

```

On the 12th January 1991 the US congress authorized the use of force to eject Iraq from Kuwait, and 'Operation Desert Storm' started on the 16th of January with US warplanes attacking targets in Iraq and Kuwait [@instituteofmedicine1997].

On that same day Fugazi played at a protest at Lafayette Park, in front of the White House:

> "The timing of this was unbelievable. Suddenly, they'd set a date to bomb Iraq, and the date we had the permit for the park was a couple of days before it," recalled Ian MacKaye, frontman of Fugazi, the headliners of the show. The city's punk dissenters arrived en masse, braving bracing temperatures. "Not only was it cold, but it was raining and snowing. It was a miserable day, and yet thousands of people came out. It was a pretty incredible show." - Ian MacKaye [@anderson2015]

The Fugazi show in front of the White House did not include either Nice New Outfit or Polish, and neither had been performed previously, suggesting that those songs probably hadn't been finished yet. It seems highly likely that the unfolding events of the US involvement in the Gulf war may have influenced those two songs, which were probably among the last songs to be written for the album.

The debut performance of Nice New Outfit was at Twisters in Richmond, VA on the 20 of February 1991 [@fugazi1991a]. The song was played a total of 114 times, which was more than Steady Diet and Polish, but less than all of the other songs on Steady Diet of Nothing.

## Lyrics

Here the lyrics of Nice New Outfit will be used as headings and commented in detail.

### You like the cut, you like the fit, wide in the shoulders, trim at the hips. What a nice new outfit, all straight clean lines.

The person who is being addressed is someone who has recently made a change, perhaps from one side to another. The critical tone of the song suggests that they might be viewed as a turncoat, someone who has shifted their allegiance for personal benefit.

### There's blood in your mouth, but not in mine.

People are going to die and it will be your responsibility not mine. This indicates disassociation between the author and the subject of the song.

### You think that I'm a sucker for it, that we're all fodder for this. Quick purchase once brought to market, dress it up and it's sold. 

This probably refers to the way the use of military force in the Persian Gulf was promoted to the American public, as if many people were going to be easily convinced. The author does not buy the arguments offered to support military intervention.

### The place I'm offered is a bedsitter's tour, a worthless sightsee of your adventure.

A bedsit is a single-room lodging so a bedsitter's tour does not go anywhere or offer anything. This conveys the frustration and powerlessness of someone railroaded into something they are opposed to - in this case, a war. "Your adventure" is similar to "your nice new outfit" and suggests bitter separation and opposition between the author and the subject of the song.

### In your nice new outfit, all straight clean lines. There's blood in your mouth, dressed to the nines. 

According to the dictionary, "dressed to the nines" means "to be wearing fashionable or formal clothes for a special occasion" [@cambridgedictionary2023]

This suggests civilian clothes, although Buick Audra commented that the person in question might be a member of the military [@wright2021].

### You're number one with a bullet, that's money well spent, 

"Number One with a Bullet" originates from the Billboard Hot 100 chart where a song at #1 but still rising was marked with a bullet symbol.

> the phrase \"number one with a bullet\" refers to a Hot 100 innovation, a small symbol indicating that a song is on its way up. To be \"number one with a bullet\" means you\'re at the top and still rising, thus not likely to vacate the penthouse anytime soon. [@giaimo2015].

Therefore, the subject is probably someone who is at the top of their area or on the rise. One possible interpretation would be that this refers to the person with the top job in the USA - the president.

"number one with a bullet" is probably also a reference to the context in which the song was probably written - the start of the gulf war - and the role that the person played in the events leading up to the war.

"that's money well spent" probably refers to donations or contributions, and strongly suggests that the subject of the song might be a politician.

### your mouth plastered like poster, address yourself success.

"your mouth plastered like poster" brings to mind TV broadcasts of speeches and interviews, specifically walls of TV screens such as those in shop windows, all the TVs tuned to the same channel, showing someone making a speech.

"address yourself success" suggests that the subject of the song might be in it for themselves, their actions motivated by personal interest and benefit. "address" is used for formal speeches, as in "address to the nation".

> When we are successful, and we will be, we have a real chance at this new world order, an order in
> which a credible United Nations can use its peace-keeping role to fulfil the promise and vision of the
> U.N.'s founders. [@bush1991]

"address yourself success" also hints at "dress for success" [@entwistle2023], which would link back to the "nice new outfit" main theme.

Perhaps there is also an echo of Dylan here, who sang on Subterranean Homesick Blues:

> Learn to dance, get dressed
>
> Get blessed, try to be a success [@bobdylan1965]

This was one of the first popular songs to be accompanied by a video (part of a documentary film), in which Dylan shows a series of prompt cards with words often varying from the lyrics - in the case of "success" the prompt card says "suck-cess" [@pennebaker1967].

### You can pinpoint your chimney and drop one down its length.

Laser-guided bombs were initially developed during the Vietnam war but their first large-scale use was during the Gulf war.

> The FLIR also contains a laser "designator" linked to a cross-hair sight on the pilot's computer display. To send a 2,000-pound bomb hurtling through the ventilation shaft of a hardened bunker, a pilot has only to head his airplane in approximately the right direction, release his bomb, fix the FLIR cross-hairs on the target, and order his laser designator to remain locked on the target. [@browne1991]

### In your nice new outfit, sorry about the mess. 

This line contrasts the "nice new outfit" with "all straight clean lines" with the impending disaster of war, in an ironic way suggesting that the subject of the song will likely keep their distance from the actual violence and will probably not accept full responsibility. "sorry about the mess" is another hint that the subject of the song is a civilian, not a member of the military, who would be unlikely to apologize, and for whom war would not be an "adventure".

## Discussion

The White House protest show and the apparent reference to laser-guided bombs in Nice New Outfit strongly suggest that both Polish and Nice New Outfit were still being worked on when the Gulf war had already started, perhaps while Fugazi were in the studio at Inner Ear recording the album. Given that the album was recorded in January 1991 this narrows down the period in which these 2 songs were probably finished to the last 19 days of January.

## Conclusion

XX

## References
