---
title: "polish with a small p"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{polish with a small p}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_polish_with_a_small_p.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

2 September 2023

## Introduction

Polish is the ninth track on Fugazi's Steady Diet of Nothing album which was recorded at Inner Ear studios in January 1991 [@fugazi1991] .

> "My journals report that I was considering the title 'Steady Diet' for a song that had the working title 'Big Choppy Chord' in late February of 1991. That song ended up being 'Polish'. In March of 1991, we decided to call the album 'Steady Diet of Nothing' and then in April we decided to include the instrumental and called that 'Steady Diet.'" [@mackaye2022]

The original title of the track was Steady Diet and it was introduced as such by Ian MacKaye the second time it was performed, at the University of Connecticut [@fugazi1991a].

> "it was i that introduced the song as 'steady diet'. it was the original title of the song (i wrote the lyrics), but we decided to use it for the instrumental that we wrote around the same time. 'steady diet of nothing' was at first just an allusion to television, but the concept grew in scope while we were writing the songs that ended up becoming the 'steady diet of nothing' album." - Ian MacKaye [@wright2021]

Judging from the dates of debut live performances in the Fugazi Live Series, the songs on Steady Diet of Nothing were written at different times in a period of about 3 years and 3 months, from October 1987 to January 1991.

```{r echo=FALSE}
releases_data_input %>%
  filter(release=="steady diet of nothing") %>%
  select(song, date) %>%
  rename(debut = date) %>%
  arrange(debut)

```

On the 12th January 1991 the US congress authorized the use of force to eject Iraq from Kuwait, and 'Operation Desert Storm' started on the 16th of January with US warplanes attacking targets in Iraq and Kuwait [@instituteofmedicine1997].

On that same day Fugazi played at a protest at Lafayette Park, in front of the White House:

> "The timing of this was unbelievable. Suddenly, they'd set a date to bomb Iraq, and the date we had the permit for the park was a couple of days before it," recalled Ian MacKaye, frontman of Fugazi, the headliners of the show. The city's punk dissenters arrived en masse, braving bracing temperatures. "Not only was it cold, but it was raining and snowing. It was a miserable day, and yet thousands of people came out. It was a pretty incredible show." - Ian MacKaye [@anderson2015]

The Fugazi show in front of the White House did not include either Nice New Outfit or Polish, and neither had been performed previously, suggesting that those songs probably hadn't been finished yet. It seems highly likely that the unfolding events of the US involvement in the Gulf war may have influenced those two songs, which were probably among the last songs to be written for the album.

Nice New Outfit contains the lines:

> You can pinpoint your chimney and drop one down its length. In your nice new outfit, sorry about the mess [@fugazi1991c]

Laser-guided bombs were initially developed during the Vietnam war but their first large-scale use was during the Gulf war [@browne1991]. The Alphabetical Fugazi episode on Nice New Outfit noted the reference to the laser-guided bombs, the context of the Gulf war, and the way the war was sold to the public [@wright2021a].

The Gulf war was nicknamed the "video-game war" due to the pervasive television coverage of bombing raids and targets apparently being hit:

> This superiority was exercised by televising the high-end war technology the US and its allies used during the war. However, in all this reporting about war, TV provided an extremely limited view of the fighting. In fact, it was a "bird-view" of the fighting as the Defense Department supplied most of the video of the air war. Recorded at night and with night vision equipment, the images resembled a video game ("Television -- The Persian Gulf War"), hence the war being nicknamed -- The Video Game War [@zigic2023].

> The uneasy relationship between war and video games has been an issue at least since the first Gulf War. At a press conference in February, 1991 General Norman Schwartzkopf felt compelled to remind Americans that "This [Operation Desert Storm] is not a video game" [@stahl2006]

> TV presented the Gulf War primarily as entertainment, complete with dramatic titles, graphics, and music. [@kellner1992]

The White House protest show and the apparent reference to laser-guided bombs in Nice New Outfit strongly suggest that both Polish and Nice New Outfit were still being worked on when the Gulf war had already started, perhaps while Fugazi were in the studio at Inner Ear recording the album. Given that the album was recorded in January 1991 this narrows down the period in which these 2 songs were probably finished to the last 19 days of January.

## Lyrics

Here the lyrics of Polish will be used as headings and commented in detail.

### Got with the program, swallowed it whole, sugar made it easy, ice made it cold.

This probably refers to the way in which television news was used by politicians and military leaders to make the case for the US going to war against Iraq and the way in which many people were easily persuaded. "Sugar" brings to mind "sugar-coated mindless garbage" from MTV-Get Off the Air [@deadkennedys1985].

"Ice" could be a seasonal reference as it was probably winter when the song was written, and perhaps a specific reference to the protest in front of the White House:

> Of the hundreds of raucous punk protest concerts organized by Positive Force, the enduring Washington DC youth activist collective, only one gave President George HW Bush insomnia and Fugazi near-frostbite. Called the Punk Percussion Protest and War On Poverty Not In the Middle East, the now-infamous demonstration was staged directly in front of the White House on 12 January 1991. Long planned as a multi-artist rally that would draw attention to homelessness in America, the event took a hairpin turn as Operation Desert Storm grew imminent. [@anderson2015]

Ice could also refer to forward-looking infrared radar (FLIR), a guidance system for bombs that requires the sensor to be very cold:

> One of the principal airborne devices used to assist attacks with Paveway-equipped bombs is the forward-looking, infrared radar, or FLIR, which is fitted to the noses of many of the American fighters and attack aircraft in the gulf region. The heart of the FLIR is a special camera with a sensor chilled to minus 320 degrees Fahrenheit. The very low temperature of this device enables it to detect the faint infrared radiation given off by warmer objects, including tanks and warm human beings. [@browne1991]

Minus 320 degrees Fahrenheit is approximately minus 195 degrees Celsius - much colder than freezing. "Ice made it cold" may also allude to the emotional detachment implied in the planning and execution of military operations.

### Reached out and touched them on a TV screen

The Gulf war was televised more than any war had been up to that time, due to the use of satellite communications.

> "We swamped the region with as many people and pieces of high technology we could afford and get permission for. It produced on-the-day war reporting on a scale never seen before. People will argue about which was the first 'television war'. Vietnam? The Israeli invasion of Lebanon? The Falklands? I have no doubt that the Gulf was the first war in which the full potential of satellite technology was used, bringing with it new technical and ethical challenges for broadcasters" [@purvis1991]

Video was used a part of the targeting systems for laser-guided missiles and this footage was re-used in press briefings.

"Reach out and touch someone" was a well-known TV commercial for the Bell and AT&T telephone companies that was used from about 1979 to 1989 [@anand2021]. Later, the phrase started being used in a darker way to refer to long-distance acts of terrorism:

> "'During the Vietnam War,' a top Bush adviser said about the menace of terrorism, 'people never thought there was the remotest chance that the North Vietnamese could actually reach out and touch someone.'" [@dowd1991]

Therefore, "reached out and touched them on a TV screen" sounds like MacKaye is using the language that the Bush administration would use to refer to terrorism to describe the actions of their own coalition in the Gulf war.

### broke out the polish, scrubbed it clean, scrubbed it clean, that dirty machine.

People use polish to make something shine - but it should be clean first. The idea of polishing seems dissonant which comes next - scrubbing clean a dirty machine. The polish is likely a visual metaphor for the blackness that characterised a lot of the videos of bombing raids that were used in press briefings by General Norman Schwarzkopf and others [@sterlingrutherford1991a; @sterlingrutherford1991b]. The videos were generally filmed from the viewpoint of either the laser designator, in which case there would be a plume of black smoke after impact, or the missile itself, in which case the screen would briefly go black after impact.

"Scrubbed it clean" could be a metaphor for a missile destroying its target. The repetition of the phrase could be for emphasis but could also imply a double meaning, bringing to mind Lady Macbeth trying to get the blood stain off her hands:

> Out, damned spot; out, I say. One, two,---why, then 'tis time to do't. Hell is murky. Fie, my lord, fie, a soldier and afeard? What need we fear who knows it when none can call our power to account? Yet who would have thought the old man to have had so much blood in him? [@shakespeare1606]

This wouldn't be the first time or the last time this Shakespeare reference appears in MacKaye's songs, there is 'Don't you know ink washes out / Easier than blood' [@fugazi1990] and also the clearest case 'Soap and water will never get rid of that spot' [@coriky2020].

"that dirty machine" suggests the machinery of war. The way in which television journalists tend to present war has been the subject of research:

> "In 'our' war coverage, television journalists tend to excise the opposite side, sanitize the suffering inflicted on it, attribute equal strength to both sides, personalize 'our' side, demonize 'their' side, and decontextualize its aggressive actions." [@liebes2009]

### This is your portion, please don't share, the only answer, your only prayer, your reflection, please don't stare, nothing is going on in...

"This is your portion" compares TV news to food and hints at how someone is deciding what gets served up to TV audiences, how much and in what way.

The list here relates to thought and perception: "the only answer, your only prayer, your reflection, please don't stare"

"Please don't share" suggests individuals are expected to receive what they are being served in isolation.

The tone and the type of message here -- a series of short statements and requests -- wrapped up in a catchy, repetitive chorus seems to allude to the type of content used in TV advertising.

The omission of the ending to the phrase "Nothing is going on in..." hints at how censorship and editing decisions may shape the perceptions of people relying on the television to inform themselves about the world.

### There is your manner, what you wear, your diseases and your repairs, all your belongings, please don't share, nothing is going on in there.

Here the list seems to refer to behaviour and needs: "your manner, what you wear, your diseases and your repairs, all your belongings"

"nothing is going on in" is repeated but this time the phrase is finished -- "nothing is going on in there". The repeated use of 'nothing', as well as being related to the original song title and the name of the album, recalls 'She Watch Channel Zero': '2, 7, 5, 4, 8, she watched, she said / All added up to zero, and nothing in her head', which featured notable use of repetition to make a point -- 'she watch' x 36 [@publicenemy1988].

### Update the cleanser, never mind the stain, we'll take the package, let the contents remain, contents remain as yet unnamed.

This seems to speak of technological change or political change linked to sanitization. Again, the image of someone trying to clean something that cannot be cleaned or is very difficult to clean. Package versus contents contrasts the superficial appearance of things, how they are presented, how they are sold, with what they really are on a deeper more intrinsic level -- which is not shared with the viewer and may not be seen at all.

### Came home flat busted only to be saved; everything comes down to you in a world concave, world concave

"Came home flat busted only to be saved" suggests someone getting home and slumping in front of the television, but that is not necessarily the only meaning, it could also refer to someone using their time and energy to do something else more worthwhile than watching television.

"everything comes down to you in a world concave" sounds like a call to action, to engage with the world and with other people -- unlike the TV screen which is convex (or it would have been at the time this song was written), we are immersed in and surrounded by the real world.

This part of the song echoes TV Party: 'We've got nothing better to do / Than watch TV and have a couple of brews' contrasted with 'Hey, wait a minute! My TV set doesn't work (It's broken!) / What are we gonna do tonight, this isn't fair! (We're hurtin'!)' [@blackflag1981].

### so well behaved, we still turn on, thirty minutes long, we still turn on

The song ends with a criticism of passivity, "so well behaved" echoing the starting line "got with the program".

"We still turn on" sounds like a weary acknowledgement that despite everything, it is hard to ignore the television coverage of the unfolding war.

"Thirty minutes long" probably refers to what used to be a typical duration of TV news broadcasts [@schultz2005].

## Discussion

It isn't always clear who is doing what in this song. Who reaches out and touches who? Who breaks out the polish? Who scrubs clean the dirty machine? Who came home flat busted, and who saved them? Some of this vagueness might be deliberate, echoing the nature of television war coverage where it isn't clear who is doing what, or what is really happening.

"You" is clearer - "everything comes down to you in a world concave" sounds like a direct appeal to the listener, perhaps echoing the sentiment of The Revolution Will Not Be Televised [@scott-heron1971].

The use of "we" seems to indicate the community formed by band and audience.

A recent discussion about this song on the podcast The Alphabetical Fugazi focused on this being a song about television:

> The stakes are higher with Reclamation and it seems like a topic that needs a plain-spoken, on-the nose statement, whereas maybe disliking television programming is not one of those topics, and that's the main difference [@wright2021].

There have been many previous songs critical of the use of television, for instance The Revolution will not be Televised [@scott-heron1971], TV Party [@blackflag1981], MTV-Get Off the Air [@deadkennedys1985], She Watch Channel Zero?! [@publicenemy1988] and (Welcome) Television [@thebeatnigs1988] -- to name some likely points of reference for the present song. What makes this song different is the focus on how television was used in the context of the build-up to the Gulf war and perhaps the early stages of the war itself. This song is also notable in how it uses a range of different literary devices: repetition, omission, euphemism, double meanings, allusions, metaphors, slang and contextual references. Even the title, 'Polish', is tricky as depending on capitalization and pronunciation it could mean two completely different things, one of which seems to be a red herring. The language used is not excessively cryptic but it isn't very direct either, compared to others such as '(Welcome) Television' [@thebeatnigs1988] which come right out and say what they want to say without requiring much effort of interpretation from the listener. The elusiveness of the lyrics brings to mind the type of language used to evade censorship, which seems appropriate to the context because during wartime the military censor all communications to and from the war zone [@omara1991].

Considering the specific context helps to make sense of the strength of feeling and urgency expressed by the song, and returning to the Fugazi show in front of the White House on the 12th of January 1991 is a good way to do this.

> The day before the event, a heavy snowstorm hit the city. The next day, temperatures nudged into the 40s and the snow melted, but is was cold and rainy. With the country preparing for war, police blanketed the White House area. Despite the weather, around 3000 people showed up, including rarely seen old allies like Squip. As the protesters beat on oil barrels, drums, tin cans, and kettles, the organizers wondered what to do about the show. Playing on an unprotected stage in the rain could expose musicians and crew to potential electrocution. At one point the rain stopped, only to begin again just as the band was ready to go. Organizers and musicians huddled on the stage, trying to decide what to do. Finally, feeling the gravity of the political moment, MacKaye said simply, 'Let's fucking do it!' All available hands scrambled to uncover the gear and prepare for the chancy performance. [@andersen2009]

The willingness of the band to play in the rain on an outdoor stage on a cold day, risking potential electrocution, clearly illustrates how strongly they felt about all that was at stake. By way of an introduction to KYEO Ian MacKaye spoke the following words:

> I gotta tell you I'm kinda terrified, it's definitely looking like a situation where you can't turn around and worst of all it's a decision that none of us can even make, except for the fact that we didn't say anything in the first place about it, so because we didn't, it's time to get even more louder and make more volume make more noise [@fugazil1991; @fugazi1991c]

The 'punk percussion protest' started at the beginning of the war and went on around the clock for several weeks at least.

> After the tocsin of war come the drums of war. "Those damned drums are keeping me up all night," the President told a gathering of Republicans on Friday. The antiwar protesters camped in Lafayette Park across the street from the White House have been beating a drum round the clock since the war began, like an Edgar Allan Poe heartbeat intended to pierce the President's conscience as he sleeps in a room above Pennsylvania Avenue [@dowd1991].

President Bush tried and failed to get the drums silenced:

> During the Gulf War, President George H.W. Bush was annoyed with an anti-war demonstration in Lafayette Park. An 'incessant' drumbeat used by protestors was so loud he ordered the drums silenced. 'Those damn drums are keeping me up all night,' Mr. Bush told a group of visiting congressmen. But a federal court held the park was a public place. The whole point of free speech was to get the attention of the President. The beat went on. - Roger McDaniel [@yesterdaysamericaeditorialteam2020]

Polish seems to be strongly related to several other songs on the Steady Diet of Nothing album, in particular Nice New Outfit and KYEO, perhaps also Stacks. Keep Your Eyes Open (KYEO) emphasises the importance of awareness complementing the concern about superficiality in Polish, with specific reference to the military: "The troops are quiet tonight, but it's not alright, because we know they're planning something". Stacks would probably also have been finished in January 1991, and several lines from this song could be interpreted as an expression of profound discontent and disagreement with the participation of the USA in the Gulf war: "I see them spinning on / So I spin out / America is just a word but I use it." The Coriky song 'Clean Kill' revisited similar themes to those explored in 'Nice New Outfit' and 'Polish', almost thirty years later [@coriky2020].

## Conclusion

This song is about the use of television in a very specific context - the participation of the US in the Gulf war (1990 - 1991). There is a lot more depth and richness to the lyrics than first impressions might suggest. The music sounds like a protest song, and the lyrics read as if they had been written to get past military censorship.

## References
