---
title: "Rebuilding the Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rebuilding the Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

## Overview

Primary data - show listings, venue coordinates, tag/duration data, and discography metadata - lives in this package's own `inst/extdata/`, not in a separate package. The companion package [`fugazi.db`](https://github.com/alexmitrani/fugazi.db) is downstream of Repeatr: it's a generated export of Repeatr's own cleaned data, not a source Repeatr reads from. See `vignette("Data-Catalogue", package = "fugazi.db")` for what it contains.

This vignette covers the three-stage process of refreshing `inst/extdata/`'s raw sources, rebuilding every dataset in `data/*.rda`, exporting the fugazi.db subset, and getting the rebuilt package and Shiny app back out into the world. `data-raw/build_data.R` is the canonical, runnable version of the process below - this vignette explains the why, that script is what to actually run.

## 1. Refreshing the raw sources (stage A)

Update whichever of `inst/extdata/`'s raw sources has new material:

- **Show data** - a fresh `scrape_fls_shows()` run, writing to `inst/extdata/fls_data.csv`.
- **Venue coordinates** - a synced export of the private Google Sheet used to look up/confirm venue locations, overwriting `inst/extdata/fls_venue_geocoding_v2.csv` directly. This file is the single source of truth for coordinates in both Repeatr and fugazi.db.
- **Tags/duration data** - re-exported from kid3 against the personally-tagged MP3 collection, overwriting `inst/extdata/fls_tags.txt`. If a new show's album string doesn't parse cleanly, the hand-written corrections for specific albums/venues live in the "process tags data" section of `R/Repeatr_1.R`.
- **Song/release/duration and tempo data** - hand-maintain `inst/extdata/releases_songs_durations_wikipedia.csv`, `inst/extdata/releases.csv`, and `inst/extdata/song_tempo_bpm_data.csv` directly against their sources (Wikipedia, rateyourmusic.com, personal BPM readings).

None of these five need refreshing together - run whichever one has new material.

## 2. Rebuilding Repeatr's own data (stage B)

Once the raw sources are refreshed, rebuild Repeatr's own data objects:

```{r}
Repeatr_Updatr(really = "really", update_stacks = TRUE)
```

This runs `Repeatr_1() -> Repeatr_2() -> Repeatr_3() -> Repeatr_4() -> Repeatr_5()` and saves every downstream dataset (`othervariables`, `Repeatr0`, `Repeatr1`, `gid_sound_quality`, `played_with`, `shows_data`, `xray`, `fls_tags`, `fls_tags_show`, `songvarslookup`, `song_tempo_bpm_data`, the choice-model outputs, and more) into `data/`. It's the `really = "not_really"` default that stops this running by accident - always pass `really = "really"` explicitly. It can take a while (the choice model fit in `Repeatr_4()` is the slow part), so it's worth checking `Repeatr_1()` on its own first if you only want to sanity-check a raw-source refresh.

`Repeatr_1()` also does two small pieces of cleaning on `othervariables`/`shows_data` beyond typed/joined raw data: it splits the raw scraped door-price text into a numeric `price` and an ISO 4217 `currency` (via `inst/extdata/fls_doorprice_currency_lookup.csv`), and it fills in Brazilian state codes for `subdivision` (the FLS site's own scrape never populates a subdivision outside the US/Canada/Australia) and standardizes any remaining blank `subdivision` to `NA`.

`update_stacks = TRUE` also regenerates `gid_initial_gid_sound_quality` (the data behind the Shiny app's "stock" pages, via `Repeatr_6()`) - `data-raw/build_data.R`'s rebuild call passes this by default, so a normal rebuild keeps it current; omit it only if you deliberately want to skip the (slower) stacks regeneration.

`Repeatr_1()` (and `Repeatr_Updatr()`, which threads it through) accepts an `output_dir` argument if you want to redirect the rebuilt `data/*.rda` objects somewhere other than the package root's own `data/` - useful for testing a rebuild without touching your working checkout.

## 3. Exporting fugazi.db (stage C)

Once Repeatr's own `data/*.rda` objects are rebuilt, compose fugazi.db's tables from them:

```{r}
export_fugazidb_data(fugazidb_dir = "../fugazi.db")
```

This reads Repeatr's own already-saved `data/*.rda` objects (no re-derivation) plus `inst/extdata/fls_venue_geocoding_v2.csv` directly, and writes `fugazidb_dir/data/*.rda` - six tables in all (`shows`, `locations`, `durations`, `discography`, `songs`, `bands`). `locations`'s coordinates are named `latitude`/`longitude`; `discography`/`songs` hold what used to be `releases`/`discography` respectively (renamed so `discography` means what it says). `fugazidb_dir` has no default - point it at a local `fugazi.db` checkout. This is a one-way generation step: it writes files, but does not commit or push anything in that checkout - review and commit fugazi.db's own changes separately, on its own schedule. Nothing in Repeatr's own pipeline depends on fugazi.db being refreshed or reinstalled.

## 4. Reinstalling and redeploying

1.  Commit and push the regenerated `data/*.rda` files.
2.  Reinstall the package from the updated source (`devtools::install()`, or `devtools::install_github("alexmitrani/Repeatr")` once pushed).
3.  Run the Shiny app locally to check nothing broke: `shiny::runApp("inst/shiny/Fugazetteer")`.
4.  Redeploy to shinyapps.io (`rsconnect::deployApp("inst/shiny/Fugazetteer")`, or the RStudio "Publish" button from `app.R`).

The app reads the package's lazy-loaded data objects for everything except venue coordinates (fetched live from a Google Sheet - see `vignette("Data-Provenance")`), so as long as stages 1-2 above ran cleanly there's nothing else to change in `app.R` itself.
