---
title: "Shows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

## Introduction

This is my list of shows I plan to go back to and listen to again once I have finished my journey through the [Fugazi Live Series](https://www.dischord.com/fugazi_live_series).

## Shows

[FLS0001, Wilson Center, Washington, DC USA 1987-09-03](https://www.dischord.com/fugazi_live_series/washington-dc-usa-90387)

[FLS0044, Maxwell's, Hoboken, NJ USA 1988-06-30](https://www.dischord.com/fugazi_live_series/hoboken-nj-usa-63088)

[FLS0095, Wilson Center, Washington, DC USA 1989-03-24](https://www.dischord.com/fugazi_live_series/washington-dc-usa-32489)

[FLS0100, Maxwell's, Hoboken, NJ USA 1989-04-09](https://www.dischord.com/fugazi_live_series/hoboken-nj-usa-40989)

[FLS0203, 2301 Canton Warehouse, Dallas, TX USA 1990-05-04](https://www.dischord.com/fugazi_live_series/dallas-tx-usa-50490)

[FLS0225, Sokol Hall, Omaha, NE USA 1990-06-10](https://www.dischord.com/fugazi_live_series/omaha-ne-usa-61090)

[FLS0230, Waukesha Expo Center, Waukesha, WI USA 1990-06-15](https://www.dischord.com/fugazi_live_series/waukesha-wi-usa-61590)

[FLS0237, Club Soda, Kalamazoo, MI USA 1990-06-24](https://www.dischord.com/fugazi_live_series/kalamazoo-mi-usa-62490)

[FLS0245, Troll, Hoorn, Netherlands 1990-09-01](https://www.dischord.com/fugazi_live_series/hoorn-netherlands-90190)

[FLS0252, Bierkeller, Bristol, England 1990-09-10](https://www.dischord.com/fugazi_live_series/bristol-england-91090)

[FLS0256, Riverside, Newcastle, England 1990-09-14](https://www.dischord.com/fugazi_live_series/newcastle-england-91490)

[FLS0262, King's Hall, Herne Bay, England 1990-09-20](https://www.dischord.com/fugazi_live_series/herne-bay-england-92090)

[FLS0272, Rock Lamour, Linkoping, Sweden 1990-10-02](https://www.dischord.com/fugazi_live_series/linkoping-sweden-100290)

[FLS0278, Glocksee, Hannover, Germany 1990-10-09](https://www.dischord.com/fugazi_live_series/hannover-germany-100990)

[FLS0287, Schlachthof, Wels, Austria 1990-10-19](https://www.dischord.com/fugazi_live_series/wels-austria-101990)

[FLS0294, Forte Prenestino, Rome, Italy 1990-10-27](https://www.dischord.com/fugazi_live_series/rome-italy-102790)

[FLS0296, Hiroshima, Torino, Italy 1990-10-29](https://www.dischord.com/fugazi_live_series/torino-italy-102990)

[FLS0303, Terminal Export, Nancy, France 1990-11-06](https://www.dischord.com/fugazi_live_series/nancy-france-110690)

[FLS0307, 9:30 Club, Washington, DC, USA 1991-01-03](https://www.dischord.com/fugazi_live_series/washington-dc-usa-10391)

[FLS0308, Lafayette Park, Washington, DC, USA 1991-01-12](https://www.dischord.com/fugazi_live_series/washington-dc-usa-11291)

[FLS0309, Sacred Heart Church, Washington, DC, USA 1991-02-15](https://www.dischord.com/fugazi_live_series/washington-dc-usa-21591)

[FLS0314, Marquee, New York, NY, USA 1991-03-05](https://www.dischord.com/fugazi_live_series/new-york-ny-usa-30591)

[FLS0315, Maxwell's, Hoboken, NJ, USA 1991-03-06](https://www.dischord.com/fugazi_live_series/hoboken-nj-usa-30691)

19910318 Church House Inn, Providence, RI, USA

19910412 Sacred Heart Church Hall, Washington, DC, USA

19910420 Columbia University, New York, NY, USA

19910502 Miracle House, Greensboro, NC, USA

19910516 Cameo Theater, Miami, FL, USA

19910518 Club Demolition, Gainesville, FL, USA

19910604 Mississippi Nights, St. Louis, MO, USA

19910617 Sacred Heart Church Hall, Washington, DC, USA

19910728 dc space, Washington, DC, USA

19910819 North Vancouver Sports Arena Hockey Rink, Vancouver, BC, Canada

19910831 Scottish Rites Temple, Oakland, CA, USA

19910901 Gilman Street, Berkeley, CA, USA

19911108 Powerstation, Auckland, New Zealand

19911112 Anti Knock, Tokyo, Japan

19911208 dc space, Washington, DC, USA

19920124 Hollywood Palladium, Los Angeles, CA, USA

19920125 Hollywood Palladium, Los Angeles, CA, USA

19920306 St. Stephen's Church Hall, Washington, DC, USA

19920413 Rockbar, New Haven, CT, USA

19920509 Brixton Academy, London, England

19920520 Vera, Gronigen, Netherlands

19920526 Boterhalle, Diksmuiden, Belgium

19920603 En Bruto, Zaragoza, Spain

19920605 Murazzi, Torino, Italy

19920613 Roxy, Prague, Czechoslovakia

19920616 Theaterfabrik, Munich, Germany

19920617 Transmitter Festival, Hohonem, Austria

19920619 Hugenotten Halle, Frankfurt, Germany

19920622 Komm, Nuremberg, Germany

19920620 Stadthalle Langenau, Ulm, Germany

19920628 Tempodrom, Berlin, Germany

19920703 Schlachthof, Bremen, Germany

19920704 Ungdomshuset, Copenhagen, Denmark

19920711 Doornroosje, Nijmegen, Netherlands

19920725 US Capitol Plaza Supreme Court, Washington, DC, USA

19921023 Sanctuary Theater, Washington, DC, USA

19921025 Steelworkers Hall, Baltimore, MD, USA

19930205 Cat's Cradle, Chapel Hill, NC, USA

19930214 Edge, Ft. Lauderdale, FL, USA

19930216 Ritz, Tampa, FL, USA

19930219 40 Watt, Athens, GA, USA

19930220 Trim Shop, Greensboro, NC, USA

19930409 New Orleans Music Hall, New Orleans, LA, USA

19930412 Liberty Lunch, Austin, TX, USA

19930413 Liberty Lunch, Austin, TX, USA

19930416 Bomb Factory, Dallas, TX, USA

19930420 Coyote's, Tucson, AZ, USA

19930414 Showcase Special Events, San Antonio, TX, USA

19930423 Hollywood Palladium, Los Angeles, CA, USA

19930425 Hollywood Palladium, Los Angeles, CA, USA

19930427 Anaconda, Santa Barbara, CA, USA

19930428 Veteran's Memorial Hall, Watsonville, CA, USA

19930502 Crest Theater, Sacramento, CA, USA

19930508 Capitol Theater, Olympia, WA, USA

19930509 YMCA, Bellevue, WA, USA

19930516 Utah State Fairpark Coliseum, Salt Lake City, UT, USA

19930524 First Avenue, Minneapolis, MN, USA

19930526 Orpheum Theater, Green Bay, WI, USA

19930528 Oak Theatre, Chicago, IL, USA

19930809 Fort Reno, Washington, DC, USA

19930816 Maxwell's, Hoboken, NJ, USA

19930817 City Gardens, Trenton, NJ, USA


