---
title: "The Emperor's New Outfit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Emperor's New Outfit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_emperors_new_outfit.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

11 September 2023

## Introduction

Nice New Outfit is the third track on Fugazi's Steady Diet of Nothing album which was recorded at Inner Ear studios in January 1991 [@fugazi1991].

Judging from the dates of debut live performances in the Fugazi Live Series, the songs on Steady Diet of Nothing were written at different times in a period of about 3 years and 3 months, from October 1987 to January 1991.

```{r echo=FALSE}
releases_data_input %>%
  filter(release=="steady diet of nothing") %>%
  select(song, date) %>%
  rename(debut = date) %>%
  arrange(debut)

```

On the 12th January 1991 the US Congress authorized the use of force to eject Iraq from Kuwait, and 'Operation Desert Storm' started on the 16th of January with US warplanes attacking targets in Iraq and Kuwait [@instituteofmedicine1997].

Fugazi played at a protest at Lafayette Park, in front of the White House, on the 12th January 1991:

> "The timing of this was unbelievable. Suddenly, they'd set a date to bomb Iraq, and the date we had the permit for the park was a couple of days before it," recalled Ian MacKaye, frontman of Fugazi, the headliners of the show. The city's punk dissenters arrived en masse, braving bracing temperatures. "Not only was it cold, but it was raining and snowing. It was a miserable day, and yet thousands of people came out. It was a pretty incredible show." - Ian MacKaye [@anderson2015]

The Fugazi show in front of the White House did not include either Nice New Outfit or Polish, and neither had been performed previously, suggesting that those songs probably hadn't been finished yet. The White House protest show and the apparent reference to laser-guided bombs in Nice New Outfit strongly suggest that both Polish and Nice New Outfit were still being worked on when the Gulf war had already started, perhaps while Fugazi were in the studio at Inner Ear recording the album. Given that the album was recorded in January 1991 this narrows down the period in which these 2 songs were probably finished to the last 19 days of January. It seems highly likely that the unfolding events of the US involvement in the Gulf war may have influenced those two songs.

> [a]s the Gulf War went well for the US, initial public skepticism became what Rollins called 'getting high on war' - a virus that infected even the punk scene. [@andersen2009]

The debut performance of Nice New Outfit was at Twisters in Richmond, VA on the 20 of February 1991 [@fugazi1991a], where there was some conflict:

> At a Fugazi show at a club in Richmond, Virginia, Positive Force activists handing out anti-war fliers were physically intimidated and nearly ejected from the premises; only the band's intervention stopped the confrontation. [@andersen2013]

The show featured some comments about the war in the interlude between And The Same and KYEO:

> Reaction... we saw it coming six months ago, hardly a reaction, saw it a year ago, saw it coming a long long time away, nothing we could do about it, yeah, I'm talking about that um, that war, you've all heard about it I'm sure. There are people who would believe, who would tell you that if you're opposed to war, then that meant that you were trying to undermine or betray the people who happen to be over there, but don't let people bullshit you, if you're opposed to the war, be opposed to the war, it's a good thing to be, who wants to be killed? - Ian MacKaye [@fugazi1991a]

Nice New Outfit was played a total of 114 times, which was more than Steady Diet and Polish, but less than all of the other songs on Steady Diet of Nothing.

Much of the song sounds like it is addressed to someone and here it is considered whether this might have been a specific person and if so, who that person might have been.

Nice New Outfit was discussed on the Alphabetical Fugazi:

> I always assumed that this song was about someone in the military - I assumed that this was a military uniform, that was my sort of going assumption, and I also want to add that I designed clothing for years, when I lived in New York, so I am also drawn to this because of my history with garment design and structure, so the fact that this song actually addresses that and addresses like cut, wide at the shoulders and trim at the hip, that's a language I speak, like fluently, but for some reason the anger attached to that set of phrases and the sentiment, the general sentiment in this song made me think that it was someone who was like a killing machine or involved in some sort of, I don't know, crimes against humanity or something - Buick Audra [@wright2021].

## Lyrics

Here the lyrics of Nice New Outfit will be used as headings and commented in detail.

### You like the cut, you like the fit, wide in the shoulders, trim at the hips. What a nice new outfit, all straight clean lines.

It sounds like the person being addressed dresses smartly and might have recently taken on a new role or made a change, perhaps from one side to another.

There was a politician who was reported to be shopping for clothes at around the time the war started:

> When the President's call came in to Representative Thomas S. Foley, Democrat of Washington, the Speaker of the House, at 5:30 P.M., Mr. Foley was shopping for shirts [@dowd1991]

However, Foley does not seem to deserve to be the subject of all the anger expressed in this song as he was consistently against the resolution on Authorization for Use of Military Force Against Iraq [@clymer1991].

### There's blood in your mouth, but not in mine.

People are going to die and it will be your responsibility not mine. This indicates disassociation between the author and the subject of the song.

### You think that I'm a sucker for it, that we're all fodder for this. Quick purchase once brought to market, dress it up and it's sold.

This probably refers to the way the use of military force in the Gulf was promoted to the American public, as if many people were going to be easily convinced. The author does not buy the arguments offered to support military intervention.

### The place I'm offered is a bedsitter's tour, a worthless sightsee of your adventure.

President W.H. Bush deployed troops in the Persian Gulf region on 9 August 1990, a week after the Iraqi invasion of Kuwait, and by the end of the year approximately 350,000 US troops had been deployed to the area - without the authorization of Congress. [@weed2015]

> In a press conference on January 9, 1991, President Bush reinforced this distinction in response to questions about the use of force resolution being debated in Congress. He was asked whether he thought he needed the resolution, and if he lost on it would he feel bound by that decision. President Bush in response stated: "I don't think I need it.... I feel that I have the authority to fully implement the United Nations resolutions." He added that he felt that he had "the constitutional authority---many attorneys having so advised me." On January 12, 1991, both houses passed the "Authorization for Use of Military Force Against Iraq Resolution" (P.L. 102-1). [@weed2015]

The use of military force was eventually authorized by Congress but it seems likely that the president would have pushed ahead with the military option even if this authorization had not been granted.

A bedsit is a single-room lodging so a bedsitter's tour does not go anywhere or offer anything. This conveys the frustration and powerlessness of someone railroaded into something they are opposed to - a war - without being offered any real choice.

"a worthless sightsee of your adventure" probably alludes to the extensive TV coverage of the war which is also explored in the song Polish [@fugazi1991b] and is discussed in more detail in "polish with a small p" [@mitrani2023].

"Your adventure" is similar to "your nice new outfit" and suggests bitter separation and opposition between the author and the subject of the song.

### In your nice new outfit, all straight clean lines. There's blood in your mouth, dressed to the nines.

According to the dictionary, "dressed to the nines" means "to be wearing fashionable or formal clothes for a special occasion" [@cambridgedictionary2023]. This suggests civilian clothes.

"all straight clean lines" indicates that everything is nice as tidy, no disorder - perhaps an allusion to lack of effective protest, dissent or obstacles in Congress despite a Democrat majority in both chambers.

### You're number one with a bullet, that's money well spent,

"Number One with a Bullet" originates from the Billboard Hot 100 chart where a song at #1 but still rising was marked with a bullet symbol.

> the phrase "number one with a bullet" refers to a Hot 100 innovation, a small symbol indicating that a song is on its way up. To be "number one with a bullet" means you're at the top and still rising, thus not likely to vacate the penthouse anytime soon. [@giaimo2015].

Therefore, the subject is probably someone who is at the top of their area or on the rise. One possible interpretation would be that this refers to the person with the top job in the USA - the president.

"number one with a bullet" is probably also a reference to the context in which the song was probably written - the start of the gulf war - and the role that the subject played in the events leading up to the war.

"that's money well spent" probably refers to donations or contributions, suggesting that the war would probably be in the interests of those who supported the subject financially. This strongly suggests that the subject of the song might be a politician.

### your mouth plastered like poster, address yourself success.

"your mouth plastered like poster" brings to mind TV broadcasts of speeches and interviews, specifically walls of TV screens such as those in shop windows, all the TVs tuned to the same channel, showing someone making a speech.

"address yourself success" suggests that the subject of the song might be in it for themselves, their actions motivated by personal interest and benefit. "address" is used for formal speeches, as in "address to the nation":

> When we are successful, and we will be, we have a real chance at this new world order, an order in which a credible United Nations can use its peace-keeping role to fulfill the promise and vision of the U.N.'s founders. [@bush1991]

"address yourself success" also hints at "dress for success" [@entwistle2023], which would link back to the "nice new outfit" theme.

Perhaps there is also an echo of Dylan here, who sang on Subterranean Homesick Blues:

> Learn to dance, get dressed Get blessed, try to be a success [@bobdylan1965]

This was one of the first popular songs to be accompanied by a video (part of a documentary film), in which Dylan shows a series of prompt cards with words often varying from the lyrics - in the case of "success" the prompt card says "suck-cess" [@pennebaker1967].

### You can pinpoint your chimney and drop one down its length.

Laser-guided bombs were initially developed during the Vietnam war but their first large-scale use was during the Gulf war.

> The FLIR also contains a laser "designator" linked to a cross-hair sight on the pilot's computer display. To send a 2,000-pound bomb hurtling through the ventilation shaft of a hardened bunker, a pilot has only to head his airplane in approximately the right direction, release his bomb, fix the FLIR cross-hairs on the target, and order his laser designator to remain locked on the target. [@browne1991]

### In your nice new outfit, sorry about the mess.

This line contrasts the "nice new outfit" with "all straight clean lines" with "mess" as a euphemism for the death and destruction of war. The apologetic tone is ironic, suggesting that the subject of the song will keep their distance from the actual violence. "sorry about the mess" is another hint that the subject of the song is a civilian, not a member of the military, who would be unlikely to apologize, and for whom war would not be an "adventure".

### In your nice new outfit, you're a fucking mess.

This final line drops the restraint and unleashes the underlying anger expressed in this song. The stark contrast between appearance (elegance) and substance (a mess) is reminiscent of The Emperor's New Clothes:

\>"But he hasn't got anything on!" said a little child. "Lor! Just hark at the innocent," said its father. And one whispered to the other what the child had said: "That little child there says he hasn't got anything on." "Why, he hasn't got anything on!" the whole crowd was shouting at last; and the Emperor's flesh crept, for it seemed to him they were right. "But all the same," he thought to himself, "I must go through with the procession." So he held himself more proudly than before, and the lords in waiting walked on bearing the train---the train that wasn't there at all. [@andersen1930]

## Discussion

If this song is about a specific individual, the person who seems to fit the description more than anyone else is George H. W. Bush, the 41st president of the US (1989 - 1993) and the commander in chief of the armed forces. There are several clues that all point in this direction:

-   "The place I'm offered is a bedsitter's tour, a worthless sightsee of your adventure" - the president seemed determined to use military force with or without authorization by Congress

-   "number one with a bullet, that's money well spent" - the presidency is the top job in the USA, and presidential campaigns require raising a lot of money which would imply future commitments to donors. President Bush had links to oil, defense, and intelligence [@baker2013], and his 1988 presidential campaign received significant support from the NRA [@kopel1996].

-   "your mouth plastered like poster, address yourself success" - seems to allude to the TV broadcast of the presidential address announcing the start of the war.

If the main subject of the song is the president, the meaning of "nice new outfit" requires some interpretation. Any president can be expected to be smartly dressed and have extensive resources and support regarding what they wear and how they present themselves in public. The president didn't get a new job or change their role significantly in the period of interest, and it seems unlikely that they would have started dressing differently. Perhaps "nice new outfit" should not be taken too literally.

"You like the cut, you like the fit, wide in the shoulders, trim at the hips" brings to mind a tailor-made suit, something that is just right for the subject of the song, just what they wanted, and makes them look good. This could be an allusion to the joint resolution that was passed on the 12th of January authorizing President Bush to use military force in the Persian Gulf. This resolution gave the president everything they needed to start the war without risking significant domestic resistance such as protests. Understood in this way, "nice new outfit" broadens the focus of the song making it critical not just of the president but also of everyone who contributed to the military action being authorized. Both House and Senate resolutions were sponsored and voted for by both Democrats and Republicans, and the Democrats had majorities in both chambers [@clymer1991]. The significant bipartisan support for the war in Congress hints at a double meaning considering another, informal meaning of "outfit" - "an organization, company, team, military unit, etc." [@cambridgedictionary2023a]

## Conclusion

Nice New Outfit is essentially an anti-war protest song, probably written at a similar time to Polish in late January 1991, when the Gulf War was unfolding. It sounds angry but elegant, it doesn't The "you" in the song is probably President George H. W. Bush. The "nice new outfit" is probably a metaphorical reference to the bi-partisan support that gave President Bush the authorization from Congress for his military "adventure" in the Gulf, despite Democrat majorities in both the House of Representatives and the Senate.

## References
