---
title: "The Tyranny of Distance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Argument}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references_the_tyranny_of_distance.bib
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

By Alex Mitrani

5 April 2026

This is going to be short and sweet because I only started listening to Tyranny of Distance this year, although I may have heard several of the songs earlier on live recordings. I like it a lot and it's grown on me a bit with each repeat listen. Plenty of good tunes that rock intensely without being abrasive or crushing. Some of the guitar lines, while being totally electric and rocking, bring to my mind more traditional instrumentation like bagpipes and fiddles, there's something folky about them.

The album title comes from a line in the song 'Six Months in a Leaky Boat' by Split Enz, a song which Ted Leo has covered twice, which itself references a history book called 'the tyranny of distance (how distance shaped Australia's history)' [@blainey1966].

> The tyranny of distance\
> Didn't stop the cavalier\
> So why should it stop me?\
> I'll conquer and stay free [@splitenz1982]

Lyrical themes are varied, several seem to have nautical themes as does the cover art, and some of them go quite deep. Some images appear in several songs, for instance 'parallel run streams' in biomusicology is followed by 'parallel or together' which makes me think of the lines of longitude which run parallel but converge at the poles. 'Dial Up' discusses over twenty years of history and the influence of british punk band Crass. Their independent way of living and working inspired many others, and they lived at a place called Dial House which was under threat of eviction at the time the music for this album was being written [@aitch2001].

The Tyranny of Distance was recorded February 17 - March 3, 2001 and it was released in June 2001 [@tedleo/pharmacists2001]. All songs written by Ted Leo who sings and plays guitar. The Pharmacists on this record included no less than 4 different drummers (James Canty, Brendan Canty, Seb Thomson and Danny Leo), James Canty (Nation of Ulysses, Make Up) on second guitar, Alex Minoff and Pete Kerlin on bass, plus Amy Dominguez on cello (who also played cello on Fugazi's The Argument a month or so earlier). Production was by Brendan Canty (Fugazi) and the recording engineer was Seb Thomson (Trans Am).

There's a very good article on the making of this record by John Vettesse that is recommended reading for further information [@vettese2016].

There is a collection of live recordings by Ted Leo and the Pharmacists (approved by the band) on the Internet Archive [@tedleoandthepharmacists2024]. The show at Brownies in New York City on 2001-09-02 is a good recording that includes live renditions of 9 of the 12 songs from Tyranny of Distance [@tedleoandthepharmacists2001].

Have a listen, it's worth at least 48 minutes and 54 seconds of your time.

## References
