---
title: "Updating the Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Updating the Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

## Overview

The package's data comes from three independent sources, updated on three different schedules:

1.  **Show data** - dates, venues, attendance, sound quality, played-with bands, notes, and tracklists, scraped from the [Fugazi Live Series](https://www.dischord.com/fugazi_live_series) website. Fully automated via `scrape_fls_shows()`.
2.  **Tags/duration data** - song-level timings, taken from my own personally-tagged MP3 collection. Manual, done outside this repository.
3.  **Venue coordinates** - latitude/longitude for each venue. Manual, looked up on Google Maps as new venues show up.

All three feed into `Repeatr_1()`, which combines them (plus a fair amount of hand-written, gid-keyed correction code accumulated over time) into the package's `data/*.rda` objects. `Repeatr_Updatr(really = "really")` runs `Repeatr_1()` and the rest of the modelling pipeline (`Repeatr_2` through `Repeatr_5`) in sequence, and the [Fugazetteer](https://alexmitrani.shinyapps.io/Fugazetteer/) Shiny app is just a consumer of those same rebuilt `data/*.rda` objects - it never reads a CSV directly, so once the package data is rebuilt and reinstalled the app picks it up automatically.

None of steps 1-3 need to happen together. Run whichever one has new material, then rebuild (step 4) and redeploy (step 5).

## 1. Updating show data

`scrape_fls_shows()` (`R/scrape_fls_shows.R`) discovers the current show listing directly from the site - including any pages added since the last run - and scrapes each show's detail page for date, venue, door price, attendance, recorded/mastered by, original source, sound quality, played-with, any official notes, and the tracklist.

Test on a small slice first:

```{r}
test_shows <- scrape_fls_shows(max_listing_pages = 1, max_shows = 3, sleepseconds = 2)
```

By default it's incremental: it only scrapes shows not already in the existing dataset (`fls_data.csv` unless you point `existing_data` elsewhere), so routine re-runs are quick. It also picks up on shows whose recording has newly gone from unavailable to available (`detect_changes`, on by default) without needing a full re-scrape - this is how ~50 shows that had gone from "no tracklist yet" to fully available got picked up in one run without touching the ~1000 shows that hadn't changed.

To actually refresh the packaged file, for real, against the live site:

```{r}
fls_data <- scrape_fls_shows(
  update_existing = TRUE,
  sleepseconds = 2,
  mycsvfilename = "inst/extdata/fls_data.csv"
)
```

A few things worth knowing:

-   It respects a delay (`sleepseconds`, default 2) between every request, listing pages and show pages alike - don't reduce this to hammer the site.
-   A full run pages through every listing page and scrapes every target show, which can be tens of minutes depending on how many shows are new. `max_shows`/`max_listing_pages` cap it for testing.
-   `fugazi-live-all-access` (FLS0000) is a standing all-access download bundle, not a show, and is always excluded automatically.
-   `fls_data.csv` supersedes the older `fugotcha.csv`, `gid_fls_id_sound_quality.csv` and `gid_fls_id_played_with.csv` - those older files are left in place for reference but nothing reads them anymore.

## 2. Updating tags/duration data

Song-level durations (`fls_tags`, `fls_tags_show`, and everything derived from them like `duration_summary`, `cumulative_duration_counts`) come from `inst/extdata/fls_tags.txt`, which isn't produced by any script - it's exported from [kid3](https://kid3.kde.org/), the audio tagger I use to tag every MP3 as I listen to it (see the [All Access](AllAccess.html) article for the listening process itself). Each show's MP3s get tagged with the track name and album set to `YYYYMMDD Venue, City, State, Country`, and duration comes along for free from the file itself.

To update: tag the new MP3s in kid3, export the tag data as a `;`-delimited text file in the same `track; artist; album; name; duration` format, and overwrite `inst/extdata/fls_tags.txt`. `fls_tags_importer()` parses it, and `Repeatr_1()` calls it automatically - no separate function to run. If a new show's `album` string doesn't parse cleanly (typos happen), `Repeatr_1()`'s "process tags data" section is where the hand-written corrections for specific albums/venues live - add a new `mutate(album = ifelse(...))` line there following the existing pattern if needed.

## 3. Updating venue coordinates

Venue coordinates are looked up by hand on Google Maps and kept in `inst/extdata/fls_venue_geocoding.csv`. `nscmov()` (`R/nscmov.R`) automates the bookkeeping around this, but not the lookup itself:

```{r}
othervariables <- nscmov()
```

This applies whatever coordinates are already confirmed in `fls_venue_geocoding.csv`, then writes `fls_venue_geocoding_update.csv` - a to-do list of every venue that's still unresolved (`checked == 0`). For each row, look the venue up on Google Maps, fill in `googlemaps_hyperlink`, `link_x`, `link_y` (and `city_disambiguation`/`guess`/`unknown` where relevant), then merge those filled-in rows back into `fls_venue_geocoding.csv` by hand. Run `nscmov()` again to confirm the new venues are picked up (`checked` should flip to 1).

## 4. Rebuilding everything

Once any of the above is updated, rebuild the package's data objects:

```{r}
Repeatr_Updatr(really = "really")
```

This runs `Repeatr_1() -> Repeatr_2() -> Repeatr_3() -> Repeatr_4() -> Repeatr_5()` and saves every downstream dataset (`othervariables`, `Repeatr0`, `Repeatr1`, `gid_sound_quality`, `played_with`, `shows_data`, `xray`, `fls_tags`, `fls_tags_show`, the choice-model outputs, and more) into `data/`. It's the `really = "not_really"` default that stops this running by accident - always pass `really = "really"` explicitly. It can take a while (the choice model fit in `Repeatr_4()` is the slow part), so it's worth checking `Repeatr_1()` on its own first if you only want to sanity-check the data-ingestion changes.

## 5. Reinstalling and redeploying

1.  Commit and push the updated `inst/extdata/*` source files and the regenerated `data/*.rda` files.
2.  Reinstall the package from the updated source (`devtools::install()`, or `devtools::install_github("alexmitrani/Repeatr")` once pushed).
3.  Run the Shiny app locally to check nothing broke: `shiny::runApp("inst/shiny/Fugazetteer")`.
4.  Redeploy to shinyapps.io (`rsconnect::deployApp("inst/shiny/Fugazetteer")`, or the RStudio "Publish" button from `app.R`).

The app only ever reads the package's lazy-loaded data objects, never the raw CSVs, so as long as steps 1-4 above ran cleanly there's nothing else to change in `app.R` itself.
