---
title: "Updating the Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Updating the Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  fig.width=7,
  fig.height=5,
  comment = "#>"
)
```

```{r, include = FALSE}
library(Repeatr)
```

## Overview

The package's data comes from five independent sources, updated on five different schedules. See `vignette("Data Provenance")` for how every dataset in `data/` maps onto these sources and the pipeline stages below.

1.  **Show data** - dates, venues, attendance, sound quality, played-with bands, notes, and tracklists, scraped from the [Fugazi Live Series](https://www.dischord.com/fugazi_live_series) website. Fully automated via `scrape_fls_shows()`.
2.  **Tags/duration data** - song-level timings, taken from my own personally-tagged MP3 collection. Manual, done outside this repository.
3.  **Venue coordinates** - latitude/longitude for each venue. Manual, looked up on Google Maps as new venues show up.
4.  **Song/release/duration data** - release, track number, instrumental/vocalist, and duration metadata per song, hand-maintained in `inst/extdata/releases_songs_durations_wikipedia.csv` against the Wikipedia Fugazi discography page. Manual, done outside this repository.
5.  **Song tempo data** - beats-per-minute per song, hand-compiled in `inst/extdata/song_tempo_bpm_data.csv`. Manual, done outside this repository; also read directly by the Shiny app.

All five feed into `Repeatr_1()` (song tempo data is read directly by the Shiny app instead), which combines them (plus a fair amount of hand-written, gid-keyed correction code accumulated over time) into the package's `data/*.rda` objects. `Repeatr_Updatr(really = "really")` runs `Repeatr_1()` and the rest of the modelling pipeline (`Repeatr_2` through `Repeatr_5`) in sequence, and the [Fugazetteer](https://alexmitrani.shinyapps.io/Fugazetteer/) Shiny app is just a consumer of those same rebuilt `data/*.rda` objects - it never reads a CSV directly (song tempo data is the one exception, read straight from `inst/extdata/song_tempo_bpm_data.csv`), so once the package data is rebuilt and reinstalled the app picks it up automatically.

None of steps 1-4 need to happen together. Run whichever one has new material, then rebuild (step 5) and redeploy (step 6).

`data-raw/build_data.R` is the canonical, runnable version of steps 1-5 below - this vignette explains the why, that script is what to actually run.

## 1. Updating show data

`scrape_fls_shows()` (`R/scrape_fls_shows.R`) discovers the current show listing directly from the site - including any pages added since the last run - and scrapes each show's detail page for date, venue, door price, attendance, recorded/mastered by, original source, sound quality, played-with, any official notes, and the tracklist. It also attaches `tour` (e.g. "1988 Fall European Tour") from the listing pages' own tour headings - that's the only field here that doesn't come from the show's own detail page, so a run with `gids` supplied directly (bypassing listing discovery) won't be able to fill it in.

Test on a small slice first:

```{r}
test_shows <- scrape_fls_shows(max_listing_pages = 1, max_shows = 3, sleepseconds = 2)
```

By default it's incremental: it only scrapes shows not already in the existing dataset (`fls_data.csv` unless you point `existing_data` elsewhere), so routine re-runs are quick. It also picks up on shows whose recording has newly gone from unavailable to available (`detect_changes`, on by default) without needing a full re-scrape - this is how ~50 shows that had gone from "no tracklist yet" to fully available got picked up in one run without touching the ~1000 shows that hadn't changed.

To actually refresh the packaged file, for real, against the live site:

```{r}
fls_data <- scrape_fls_shows(
  update_existing = TRUE,
  sleepseconds = 2,
  mycsvfilename = "inst/extdata/fls_data.csv"
)
```

A few things worth knowing:

-   It respects a delay (`sleepseconds`, default 2) between every request, listing pages and show pages alike - don't reduce this to hammer the site.
-   A full run pages through every listing page and scrapes every target show, which can be tens of minutes depending on how many shows are new. `max_shows`/`max_listing_pages` cap it for testing.
-   `fugazi-live-all-access` (FLS0000) is a standing all-access download bundle, not a show, and is always excluded automatically.
-   `fls_data.csv` supersedes the older `fugotcha.csv`, `gid_fls_id_sound_quality.csv` and `gid_fls_id_played_with.csv` - those older files are left in place for reference but nothing reads them anymore.
-   If `fls_data.csv` ever needs `tour`/`city`/`state`/`country` backfilled without a full show-by-show re-scrape (e.g. after adding one of these columns, or if some rows are missing it), `scrape_fls_listing_data()` crawls just the listing pages - a few dozen requests instead of 1000+ - and returns a `gid` + all four: `fls_data <- fls_data %>% select(-tour) %>% left_join(scrape_fls_listing_data() %>% select(gid, tour), by = "gid")`. `Repeatr_1()` reads `tour` straight from `fls_data.csv`/`Repeatr0` now, not from the older `fugazi-small.csv` file - `city`/`state`/`country` are scraped the same way but haven't been migrated over to source from `fls_data.csv` yet.

## 2. Updating tags/duration data

Song-level durations (`fls_tags`, `fls_tags_show`, and everything derived from them like `duration_summary`, `cumulative_duration_counts`) come from `inst/extdata/fls_tags.txt`, which isn't produced by any script - it's exported from the mp3 files using [kid3](https://kid3.kde.org/). Each show's MP3s get tagged with the track name and album set to `YYYYMMDD Venue, City, State, Country`, and duration comes along for free from the file itself.

To update: open the new mp3 files in kid3, export the tag data as format "CSV quoted" with header `track; artist; album; name; duration` and tracks `"%{track}";"%{artist}";"%{album}";"%{title}";"%{duration}"`. Open the resulting file in a text editor, remove the quotation marks and add spaces after the semi-colons using find and replace, then and add the new data to the end of any existing `inst/extdata/fls_tags.txt` file, or create the file if it does not exist. **Save as UTF-8**, not the editor's default ANSI/system codepage - a venue or city name with an accent (é, ö, ã, ...) saved in the wrong encoding will parse fine at first but crash `Repeatr_1()` with an `invalid multibyte string` error once it reaches `nchar()`/`gregexpr()` further down. `fls_tags_importer()` parses it, and `Repeatr_1()` calls it automatically - no separate function to run. If a new show's `album` string doesn't parse cleanly (typos happen), `Repeatr_1()`'s "process tags data" section is where the hand-written corrections for specific albums/venues live - add a new `mutate(album = ifelse(...))` line there following the existing pattern if needed.

## 3. Updating venue coordinates

Venue coordinates now come primarily from `inst/extdata/fls_venue_geocoding_v2.csv`, a local snapshot of a private Google Sheet the coordinates are actually maintained in - `Repeatr_1()` reads this file directly (matched to shows by `country`/`city`/`venue`), not the live sheet itself, since that sheet isn't reliably available and isn't part of the formal package workflow. To update: look up new/corrected venues on Google Maps in the sheet as usual, then download it and overwrite `inst/extdata/fls_venue_geocoding_v2.csv` with the current export.

`inst/extdata/fugazi-small.csv` is still consulted as a fallback for any venue `fls_venue_geocoding_v2.csv` doesn't cover, so a show doesn't lose its coordinates entirely just because a new venue hasn't been added to the sheet yet.

The older `fls_venue_geocoding.csv` file and `nscmov()`'s (`R/nscmov.R`) to-do-list bookkeeping around it are **no longer part of this pipeline** - `Repeatr_1()` stopped reading that file once `fls_venue_geocoding_v2.csv` became the source of truth. They're left in place for reference but nothing in `Repeatr_1()` calls `nscmov()` or reads `fls_venue_geocoding.csv` anymore; treat that workflow as retired unless it gets revived deliberately.

## 4. Updating song/release/duration and tempo data

`inst/extdata/releases_songs_durations_wikipedia.csv` carries release, track number, instrumental/vocalist, and duration metadata per song, hand-maintained against the [Wikipedia Fugazi discography page](https://web.archive.org/web/20201112000517/http://en.wikipedia.org/wiki/Fugazi_discography). `Repeatr_1()` reads it into `songvarslookup` and joins it onto the live, classified song set by song title (`song`) - **not** by a hardcoded id column, precisely so this file can't silently drift out of sync with `Repeatr_1()`'s classification rules the way it once did. If you add, rename, or remove a song here, `Repeatr_1()` will `warning()` at build time listing any song names that don't match between the live classification and this file - see the songid reconciliation check in `R/Repeatr_1.R` for what that looks like, and fix the mismatch in this CSV's `song` column before trusting the rebuilt data.

`inst/extdata/song_tempo_bpm_data.csv` carries hand-compiled tempo (BPM) per song. Unlike the other four sources, it is **not** read by `Repeatr_1()` at all - the Shiny app (`inst/shiny/Fugazetteer/app.R`) reads it directly - so there is no automated check tying its song names to the live classification; when a song's title changes elsewhere, update it here by hand too.

## 5. Rebuilding everything

Once any of sources 1-4 is updated, rebuild the package's data objects:

```{r}
Repeatr_Updatr(really = "really", update_stacks = TRUE)
```

This runs `Repeatr_1() -> Repeatr_2() -> Repeatr_3() -> Repeatr_4() -> Repeatr_5()` and saves every downstream dataset (`othervariables`, `Repeatr0`, `Repeatr1`, `gid_sound_quality`, `played_with`, `shows_data`, `xray`, `fls_tags`, `fls_tags_show`, the choice-model outputs, and more) into `data/`. It's the `really = "not_really"` default that stops this running by accident - always pass `really = "really"` explicitly. It can take a while (the choice model fit in `Repeatr_4()` is the slow part), so it's worth checking `Repeatr_1()` on its own first if you only want to sanity-check the data-ingestion changes.

`update_stacks = TRUE` also regenerates `gid_initial_gid_sound_quality` (the data behind the Shiny app's "stock" pages, via `Repeatr_6()`) - `data-raw/build_data.R`'s rebuild call passes this by default, so a normal rebuild keeps it current; omit it only if you deliberately want to skip the (slower) stacks regeneration.

## 6. Reinstalling and redeploying

1.  Commit and push the updated `inst/extdata/*` source files and the regenerated `data/*.rda` files.
2.  Reinstall the package from the updated source (`devtools::install()`, or `devtools::install_github("alexmitrani/Repeatr")` once pushed).
3.  Run the Shiny app locally to check nothing broke: `shiny::runApp("inst/shiny/Fugazetteer")`.
4.  Redeploy to shinyapps.io (`rsconnect::deployApp("inst/shiny/Fugazetteer")`, or the RStudio "Publish" button from `app.R`).

The app reads the package's lazy-loaded data objects for everything except `song_tempo_bpm_data` (read directly from its source CSV, see step 4), so as long as steps 1-5 above ran cleanly there's nothing else to change in `app.R` itself.
